### 2011-10-11 paper.tex
initial commit
template for a paper, basically copying the itp2011 qbf-validity paper.

\documentclass{llncs}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{xspace}
\usepackage{color}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\usepackage[
  colorlinks=true,
  linkcolor=linkcolor,
  citecolor=linkcolor,
  filecolor=linkcolor,
  urlcolor=linkcolor]{hyperref}

\makeatletter
\providecommand*{\toclevel@title}{0}
\providecommand*{\toclevel@author}{1}
\makeatother

\hypersetup{bookmarksdepth=subsection}

\newcommand{\HOL}{HOL4\xspace}
\newcommand{\Squolem}{Squolem\xspace}
\newcommand{\QBV}{QBV\xspace}
\newcommand{\Minisat}{MiniSat\xspace}
\newcommand{\OpenTheory}{OpenTheory\xspace}

\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\thmref}[1]{Theorem~\ref{#1}}
\newcommand{\eqnref}[1]{(\ref{#1})}

\newcommand{\sem}[1]{[\![{#1}]\!]}

\newcommand{\ie}{i.e., }
\newcommand{\eg}{e.g., }
\newcommand{\cf}{cf.\xspace}
\newcommand{\wrt}{wrt.\xspace}

\newcommand{\ite}[3]{\ensuremath{\mathsf{if}\ #1\ \mathsf{then}\ #2\ \mathsf{else}\ #3}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\hypersetup{
  pdfauthor = {Ramana Kumar and Joe Hurd},
  pdftitle = {Standalone Tactics using OpenTheory},
  pdfkeywords = {Interactive Theorem Proving, LCF-style proof checking, HOL, \OpenTheory},
  pdfsubject = {}
}

\title{Standalone Tactics using OpenTheory}

\author{Ramana Kumar\thanks{supported by the Gates Cambridge Trust}\and Joe Hurd}

\institute{\email{rk436@cam.ac.uk} and \email{joe@galois.com}}

\maketitle

\begin{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}
\label{sec:related-work}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Background and Theory}
\label{sec:background}

\subsection{Quantified Boolean Formulae}

\subsection{Skolem Functions and Models}
\label{sec:models}

\subsection{Certificates of Validity}
\label{sec:certificates}

\subsection{Higher-Order Logic}
\label{sec:hol}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Implementation}
\label{sec:checking}

\subsection{Overview}
\label{sec:checking-overview}

\subsection{Subsection}
\label{sec:preproc}

\subsection{Subsection}
\label{sec:hypelim}

\subsection{Example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Experimental Results}
\label{sec:experimental-results}

\subsection{Run-Times}
\label{sec:run-times}

\subsection{Profiling and Future Improvements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bigskip

\noindent \textbf{Acknowledgments} The authors would like to thank ....

\vspace{-\bigskipamount}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{splncs}
\bibliography{paper}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2011-10-14 paper.tex
first stab at abstract

@@
 \begin{abstract}
+Tactics, functions that compute low-level inferences to achieve a high-level reasoning goal, are usually developed within and tied to a particular proof assistant, which leads to code duplication across different systems and inhibits reuse.
+As formal methods are applied to larger problems, projects spanning multiple proof assistants benefit more from interoperability and access to the union of all implemented tactics.
+Using the OpenTheory proof exchange format, we show how to turn a tactic implemented for one system into one available to many via the web.
+This enables, for example, LCF-style proof reconstruction efforts for fast automatic provers to be shared by users of multiple proof assistants and removes the need for each user to install the automatic prover.
 \end{abstract}

### 2011-10-14 paper.tex
write paper outline and clean up tex file

@@
 
-\usepackage{amsmath}
-\usepackage{amssymb}
 \usepackage{url}
 \usepackage{xspace}
-\usepackage{color}
-\definecolor{linkcolor}{rgb}{0,0,0.5}
-\usepackage[
-  colorlinks=true,
-  linkcolor=linkcolor,
-  citecolor=linkcolor,
-  filecolor=linkcolor,
-  urlcolor=linkcolor]{hyperref}
 
-\makeatletter
-\providecommand*{\toclevel@title}{0}
-\providecommand*{\toclevel@author}{1}
-\makeatother
-
-\hypersetup{bookmarksdepth=subsection}
-
-\newcommand{\HOL}{HOL4\xspace}
-\newcommand{\Squolem}{Squolem\xspace}
-\newcommand{\QBV}{QBV\xspace}
-\newcommand{\Minisat}{MiniSat\xspace}
 \newcommand{\OpenTheory}{OpenTheory\xspace}
-
-\newcommand{\secref}[1]{Section~\ref{#1}}
-\newcommand{\figref}[1]{Figure~\ref{#1}}
-\newcommand{\thmref}[1]{Theorem~\ref{#1}}
-\newcommand{\eqnref}[1]{(\ref{#1})}
-
-\newcommand{\sem}[1]{[\![{#1}]\!]}
-
-\newcommand{\ie}{i.e., }
-\newcommand{\eg}{e.g., }
-\newcommand{\cf}{cf.\xspace}
-\newcommand{\wrt}{wrt.\xspace}
-
-\newcommand{\ite}[3]{\ensuremath{\mathsf{if}\ #1\ \mathsf{then}\ #2\ \mathsf{else}\ #3}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\hypersetup{
-  pdfauthor = {Ramana Kumar and Joe Hurd},
-  pdftitle = {Standalone Tactics using OpenTheory},
-  pdfkeywords = {Interactive Theorem Proving, LCF-style proof checking, HOL, \OpenTheory},
-  pdfsubject = {}
-}
-
 \title{Standalone Tactics using OpenTheory}
 
-\author{Ramana Kumar\thanks{supported by the Gates Cambridge Trust}\and Joe Hurd}
+\author{Ramana Kumar\thanks{supported by the Gates Cambridge Trust}\inst{1}\and Joe Hurd\inst{2}}
 
-\institute{\email{rk436@cam.ac.uk} and \email{joe@galois.com}}
+\institute{\email{Ramana.Kumar@cl.cam.ac.uk}\and\email{joe@galois.com}}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
 
-\section{Introduction}
-\label{sec:introduction}
+\section{Tactics in Proof Assistants}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Examples of proof assistants.
+Examples of tactics.
+Examples of duplication.
+
+\section{\OpenTheory for Tactic Communication}
+
+Introduction to article format.
+Representing terms and proofs in the format.
+Viewing tactics as derived inference rules.
+
+\section{Making Existing Tactics into Web Services}
+
+Implementation details to make standalone.
+Implementation details to put on the web.
+Examples of tactics for which it is done: HolQbf, HolSmt, HolSat, sledgehammer, TFL, function package, Metis, datatype packages, arithmetic decision procedure.
+
+\section{Performance}
+
+Speed comparisons against using built-in tactics.
+Amount of code and time required, compared to porting tactics at the implementation level.
 
 \section{Related Work}
-\label{sec:related-work}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Background and Theory}
-\label{sec:background}
-
-\subsection{Quantified Boolean Formulae}
-
-\subsection{Skolem Functions and Models}
-\label{sec:models}
-
-\subsection{Certificates of Validity}
-\label{sec:certificates}
-
-\subsection{Higher-Order Logic}
-\label{sec:hol}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Implementation}
-\label{sec:checking}
-
-\subsection{Overview}
-\label{sec:checking-overview}
-
-\subsection{Subsection}
-\label{sec:preproc}
-
-\subsection{Subsection}
-\label{sec:hypelim}
-
-\subsection{Example}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experimental Results}
-\label{sec:experimental-results}
-
-\subsection{Run-Times}
-\label{sec:run-times}
-
-\subsection{Profiling and Future Improvements}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Evidential tool bus.
+System on TPTP.
+Natural extension of the integration of automatic tools into interactive systems.
 
 \section{Conclusions}
-\label{sec:conclusion}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bigskip
-
-\noindent \textbf{Acknowledgments} The authors would like to thank ....
-
-\vspace{-\bigskipamount}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Additional (non-proof) work tactics might also do needs to be represented on another channel.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2011-10-18 paper.tex
start introduction, clean up and add to bib file

@@
 \newcommand{\OpenTheory}{OpenTheory\xspace}
+\newcommand{\secref}[1]{Section~\ref{sec:#1}}
 
@@
 \section{Introduction}
+There are many systems for interactively developing machine-checked formal theories, including HOL4~\cite{slind08brief}, HOL Light~\cite{DBLP:conf/tphol/Harrison09a}, ProofPower~\cite{ProofPower}, Isabelle/HOL~\cite{wenzel08isabelle}, and Coq~\cite{bertot08short}.
+%All but the last are based on higher-order logic~\cite{church40formulation,gordon94hol} and follow the LCF tradition~\cite{gordon79lcf} wherein theorems are represented by an abstract datatype all of whose constructors, representing inference rules, are contained in a small, trusted kernel.
+%In LCF-style provers, powerful proof procedures are kept trustworthy by, ultimately, only calling kernel functions to create theorems.
+The logic implemented by these systems (except Coq) is essentially the same, but the collections of libraries and tactics built atop the logical kernels differ.
+Where similar functionality exists in two systems it is usually the result of duplicate effort.
+For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into a higher order logic based system (HOL4 in one case and HOL Light in the other).
+
+We contend that tactics or libraries for proof assistants need only be written and maintained in one place rather than once per system. 
+An added advantage when the tactic is an integration of an external tool is that a user of the proof assistant need not also have installed the external tool, because a standalone tactic can be accessed from a web server where the tool is installed.
+The contributions of this paper are:
+\begin{itemize}
+\item
+A tested methodology (\secref{opentheory}) for making existing libraries standalone, using the \OpenTheory~\cite{hurd2009} article format;
+\item
+A variety of standalone tactics ready for use as web services (\secref{web}); and,
+\item
+Performance data (\secref{performance}) comparing standalone tactics against direct implementations.
+\end{itemize}
 
@@
 
-Examples of proof assistants.
-Examples of tactics.
-Examples of duplication.
+Classification and examples of the kinds of libraries and tactics implemented ontop of the kernel in various systems.
+
+Certain kinds of functionality, or improvements in functionality (e.g. Isabelle function package over TFL) become restricted to a single proof assistant unnecessarily.
 
 \section{\OpenTheory for Tactic Communication}
+\label{sec:opentheory}
 
@@
 Viewing tactics as derived inference rules.
+Implementation details to make standalone.
 
 \section{Making Existing Tactics into Web Services}
+\label{sec:web}
 
-Implementation details to make standalone.
 Implementation details to put on the web.
@@
 \section{Performance}
+\label{sec:performance}
 

### 2011-10-20 paper.tex
Weber and Amjad as an example of multiple integrations

@@
 For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into a higher order logic based system (HOL4 in one case and HOL Light in the other).
+Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems (HOL4, HOL Light, and Isabelle/HOL); each integration requires a separate implementation.
 
@@
 
+Highly customised tactics for a particular proof system might suffer performance hits if the factorisation through the OpenTheory kernel cannot be done efficiently.
+By the same token, the article format might admit optimisation by restructuring low-level proofs, and such optimisers could be widely applied.
+
 \bibliographystyle{splncs}

### 2012-01-15 paper.tex
slight reorganisation, towards rough diamond form

@@
 
-\section{Tactics in Proof Assistants}
+OpenTheory is designed for storing and sharing higher-order logic theories.
+By using OpenTheory to communicate to tactics (sending goals and receiving proofs), we obtain
+\begin{enumerate}
+\item
+A standard format (article files) for goals and proofs, and
+\item
+An extensible way to fix the meaning of constants and type definitions.
+\end{enumerate}
+
+\section{Making Existing Tactics into Web Services}
+\label{sec:web}
+
+\subsection{Tactics in Proof Assistants}
 
@@
 
+Examples of tactics: HolQbf, HolSmt, HolSat, sledgehammer, TFL, function package, Metis, datatype packages, arithmetic decision procedure.
+
 Certain kinds of functionality, or improvements in functionality (e.g. Isabelle function package over TFL) become restricted to a single proof assistant unnecessarily.
 
-\section{\OpenTheory for Tactic Communication}
+\subsection{\OpenTheory for Tactic Communication}
 \label{sec:opentheory}
@@
 Implementation details to make standalone.
+Implementation details to put on the web.
 
-\section{Making Existing Tactics into Web Services}
-\label{sec:web}
-
-Implementation details to put on the web.
-Examples of tactics for which it is done: HolQbf, HolSmt, HolSat, sledgehammer, TFL, function package, Metis, datatype packages, arithmetic decision procedure.
-
-\section{Performance}
+\section{Preliminary Performance Results}
 \label{sec:performance}
@@
 System on TPTP.
+Omega.
 Natural extension of the integration of automatic tools into interactive systems.

### 2012-01-16 paper.tex
a bit more reorganisation and planning

@@
 Where similar functionality exists in two systems it is usually the result of duplicate effort.
-For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into a higher order logic based system (HOL4 in one case and HOL Light in the other).
+
+For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into a higher-order logic based system (HOL4 in one case and HOL Light in the other).
 Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems (HOL4, HOL Light, and Isabelle/HOL); each integration requires a separate implementation.
+<Metis integrated many times separately; currenty HOL4 version is very old and upgrading would be painful>
+<Sledgehammer <ref> regarded as one of the most successful integrations <ref> is available for Isabelle only, but the ideas behind it don't intrinsically exclude HOL4, HOL Light, etc.>
+<TFL, function package,>
+Certain kinds of functionality, or improvements in functionality (e.g. Isabelle function package over TFL) become restricted to a single proof assistant unnecessarily.
+<datatype packages>
+<arithmetic decision procedures>
 
@@
 An added advantage when the tactic is an integration of an external tool is that a user of the proof assistant need not also have installed the external tool, because a standalone tactic can be accessed from a web server where the tool is installed.
-The contributions of this paper are:
-\begin{itemize}
-\item
-A tested methodology (\secref{opentheory}) for making existing libraries standalone, using the \OpenTheory~\cite{hurd2009} article format;
-\item
-A variety of standalone tactics ready for use as web services (\secref{web}); and,
-\item
-Performance data (\secref{performance}) comparing standalone tactics against direct implementations.
-\end{itemize}
 
-OpenTheory is designed for storing and sharing higher-order logic theories.
-By using OpenTheory to communicate to tactics (sending goals and receiving proofs), we obtain
+\OpenTheory~\cite{hurd2009} is designed for storing and sharing higher-order logic theories.
+By using \OpenTheory to communicate to tactics (sending goals and receiving proofs), we obtain
 \begin{enumerate}
@@
 
+The contributions of this rough diamond are:
+\begin{itemize}
+\item
+A method (\secref{implementation}) for making existing libraries standalone, using the \OpenTheory article format, including working examples online; and,
+\item
+Preliminary performance data (\secref{performance}) comparing standalone tactics against direct implementations.
+\end{itemize}
+
 \section{Making Existing Tactics into Web Services}
-\label{sec:web}
-
-\subsection{Tactics in Proof Assistants}
-
-Classification and examples of the kinds of libraries and tactics implemented ontop of the kernel in various systems.
-
-Examples of tactics: HolQbf, HolSmt, HolSat, sledgehammer, TFL, function package, Metis, datatype packages, arithmetic decision procedure.
-
-Certain kinds of functionality, or improvements in functionality (e.g. Isabelle function package over TFL) become restricted to a single proof assistant unnecessarily.
+\label{sec:implementation}
 
 \subsection{\OpenTheory for Tactic Communication}
-\label{sec:opentheory}
-
 Introduction to article format.
@@
 Viewing tactics as derived inference rules.
+
+\subsection{Isolating tactics from their home systems}
 Implementation details to make standalone.
+<requires opentheory reader and writer for the system.
+so far accomplished by instrumenting the kernel to log primitive inferences.
+this is work, but has the additional benefit of allowing import/export of theories in opentheory repositories>
+<either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>
+
 Implementation details to put on the web.
+<simple cgi - worth mentioning?>
 
@@
 Speed comparisons against using built-in tactics.
+<compare builtin holqbf, holqbf with builtin minisat, and holqbf with external minisat (i.e. chain of two webtactics) on a simple and difficult qbf problem>
+
 Amount of code and time required, compared to porting tactics at the implementation level.

### 2012-01-17 paper.tex
some references

@@
 <Metis integrated many times separately; currenty HOL4 version is very old and upgrading would be painful>
-<Sledgehammer <ref> regarded as one of the most successful integrations <ref> is available for Isabelle only, but the ideas behind it don't intrinsically exclude HOL4, HOL Light, etc.>
+<Sledgehammer~\cite{DBLP:journals/japll/MengP09,Paulson_threeyears} is available for Isabelle only, but the ideas behind it don't intrinsically exclude HOL4, HOL Light, etc.>
 <TFL, function package,>
@@
 
-Evidential tool bus.
-System on TPTP.
+System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}.
+<Comparison to THF?>
+Evidential tool bus~\cite{DBLP:conf/icfem/Rushby05,fk:etb}.
 Omega.

### 2012-01-24 paper.tex
reworking the introduction
move main novelty further up. split examples into categories.

@@
 
-For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into a higher-order logic based system (HOL4 in one case and HOL Light in the other).
-Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems (HOL4, HOL Light, and Isabelle/HOL); each integration requires a separate implementation.
-<Metis integrated many times separately; currenty HOL4 version is very old and upgrading would be painful>
-<Sledgehammer~\cite{DBLP:journals/japll/MengP09,Paulson_threeyears} is available for Isabelle only, but the ideas behind it don't intrinsically exclude HOL4, HOL Light, etc.>
-<TFL, function package,>
-Certain kinds of functionality, or improvements in functionality (e.g. Isabelle function package over TFL) become restricted to a single proof assistant unnecessarily.
-<datatype packages>
-<arithmetic decision procedures>
-
-We contend that tactics or libraries for proof assistants need only be written and maintained in one place rather than once per system. 
-An added advantage when the tactic is an integration of an external tool is that a user of the proof assistant need not also have installed the external tool, because a standalone tactic can be accessed from a web server where the tool is installed.
-
-\OpenTheory~\cite{hurd2009} is designed for storing and sharing higher-order logic theories.
-By using \OpenTheory to communicate to tactics (sending goals and receiving proofs), we obtain
+It makes sense to speak of similar functionality in different systems not just because they use roughly the same logic, but also because there is a shared base of concepts: booleans, inductive datatypes, recursive functions, natural numbers, lists, sets, etc.
+The \OpenTheory standard library~\cite{DBLP:conf/nfm/Hurd11} formalises this shared base, using \OpenTheory~\cite{hurd2009}, a format designed for storing and sharing higher-order logic theories.
+We use \OpenTheory to communicate to tactics (sending goals and receiving proofs), and thereby obtain
 \begin{enumerate}
@@
 \end{enumerate}
+
+Examples of duplicate effort include the integration of external tools into HOL-based provers.
+For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into a higher-order logic based system (HOL4 in one case and HOL Light in the other).
+Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems (HOL4, HOL Light, and Isabelle/HOL); each integration requires a separate implementation.
+Sledgehammer~\cite{DBLP:journals/japll/MengP09,Paulson_threeyears} is only available for Isabelle/HOL, but would be useful if available for other provers.
+
+Besides duplicate initial effort, costs of maintenance can be multiplied over different provers, and improvements in functionality can become restricted to single prover unnecessarily.
+For instance, Hurd's first-order prover, Metis~\cite{DBLP:conf/cade/Hurd02}, is integrated in at least three HOL-based provers, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
+Another example is a package for defining recursive functions, TFL <ref>, which was originally written within both Isabelle/HOL and HOL4 <and others?> but was recently superceded by Krauss's function package <ref> in Isabelle/HOL only.
+Krauss's improvements ought to be applicable to other HOL-based provers <really?>, but a direct re-implementation would require substantial effort.
+
+We contend that tactics or libraries for proof assistants need only be written and maintained in one place rather than once per system. 
+An added advantage when the tactic is an integration of an external tool is that a user of the proof assistant need not also have installed the external tool, because a standalone tactic can be accessed from a web server where the tool is installed.
 
@@
 \section{Conclusions}
-
 Additional (non-proof) work tactics might also do needs to be represented on another channel.
@@
 
+Other examples of functionality that could usefully be made prover-independent include packages for defining recursive datatypes or for solving classes of arithmetic problems.
+
 \bibliographystyle{splncs}

### 2012-01-24 paper.tex
a few bits and pieces

@@
 Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems (HOL4, HOL Light, and Isabelle/HOL); each integration requires a separate implementation.
-Sledgehammer~\cite{DBLP:journals/japll/MengP09,Paulson_threeyears} is only available for Isabelle/HOL, but would be useful if available for other provers.
+Sledgehammer~\cite{DBLP:journals/japll/MengP09,Paulson_threeyears,DBLP:conf/frocos/BlanchetteBN11} is only available for Isabelle/HOL, but would be useful if available for other provers.
 
@@
 For instance, Hurd's first-order prover, Metis~\cite{DBLP:conf/cade/Hurd02}, is integrated in at least three HOL-based provers, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
-Another example is a package for defining recursive functions, TFL <ref>, which was originally written within both Isabelle/HOL and HOL4 <and others?> but was recently superceded by Krauss's function package <ref> in Isabelle/HOL only.
-Krauss's improvements ought to be applicable to other HOL-based provers <really?>, but a direct re-implementation would require substantial effort.
+Another example is a package for defining recursive functions, TFL~\cite{slindphd}, originally implemented for both Isabelle/HOL and HOL4, but recently superceded in Isabelle/HOL only by Krauss's function definition package~\cite{DBLP:journals/jar/Krauss10}.
+The improvements of Krauss's method over TFL ought to be applicable to other HOL-based provers, but a direct re-implementation would require substantial effort.
 
@@
 
+<extensions to other logics?>
+
 \bibliographystyle{splncs}

### 2012-01-26 paper.tex
minor tweaks on intro

@@
 The \OpenTheory standard library~\cite{DBLP:conf/nfm/Hurd11} formalises this shared base, using \OpenTheory~\cite{hurd2009}, a format designed for storing and sharing higher-order logic theories.
-We use \OpenTheory to communicate to tactics (sending goals and receiving proofs), and thereby obtain
+We use \OpenTheory to communicate with tactics (sending goals and receiving proofs), and thereby obtain
 \begin{enumerate}
 \item
-A standard format (article files) for goals and proofs, and
+A standard format (\OpenTheory article files) for goals and proofs, and
 \item
@@
 
-Examples of duplicate effort include the integration of external tools into HOL-based provers.
+Examples of duplicated effort on tactics include the integration of external tools into HOL-based provers.
 For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into a higher-order logic based system (HOL4 in one case and HOL Light in the other).

### 2012-01-26 paper.tex
start work on section 2.1

@@
 
-\usepackage{url}
-\usepackage{xspace}
+\usepackage{url,xspace,amssymb}
 
@@
 \subsection{\OpenTheory for Tactic Communication}
-Introduction to article format.
-Representing terms and proofs in the format.
-Viewing tactics as derived inference rules.
+The central concept in \OpenTheory is that of a \emph{theory package}, $\Gamma\rhd\Delta$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
+The basic unit\footnote{composite theory packages are also possible} for constructing theory packages is an \emph{article file}, which consists of instructions for a virtual machine whose operations include primitive inference rules of higher-order logic and basic term-manipulation.
+
+<Intro to LCF kernel? Viewing tactics as derived inference rules?>
+
+An article file represents a theory $\Gamma\rhd\Delta$.
+As a special case, we can represent the proof of a single goal $g$ by taking $\Delta=\{\vdash g\}$.
+More generally, by taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
+
+Furthermore, we can represent a higher-order logic term $t$ by the theory $\{\vdash\mathtt{K\;F\;}t\}\rhd\{\vdash\mathtt{K\;F\;}t\}$, where $\mathtt{K}:\alpha\to\beta\to\alpha$ is the combinator defined by $\mathtt{K}\;x\;y=x$, and $\mathtt{F}$ is falsity (any boolean would do).
+The article format ensures $t$ is only written once in representing this theory.
+This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.
+
+Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
+(Additional inputs and outputs can always be communicated separately.)
 

### 2012-01-26 paper.tex
start on 2.2

@@
 
-<Intro to LCF kernel? Viewing tactics as derived inference rules?>
+<more concrete examples? definition of tactics? viewing tactic as derived inference rule? intro to LCF kernel?>
 
@@
 \subsection{Isolating tactics from their home systems}
-Implementation details to make standalone.
-<requires opentheory reader and writer for the system.
-so far accomplished by instrumenting the kernel to log primitive inferences.
-this is work, but has the additional benefit of allowing import/export of theories in opentheory repositories>
+There are two approaches to obtaining a standalone tactic: either write one directly, or extract an existing tactic from a proof assistant. 
+We have experimented with the second approach, taking tactics from HOL4 and from HOL Light.
+<The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.>
+The key requirements of the proof assistant are:
+\begin{enumerate}
+\item
+An \OpenTheory article file reader and writer,
+\item
+A facility for recording proofs and converting them to use the \OpenTheory primitive inferences, and,
+\item
+A facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
+\end{enumerate}
+<first two items are the most work, but have the additional benefit of allowing import/export of theories in opentheory repositories>
+<in HOL4 and HOL Light, item 2 is accomplished by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
+this step is made easy because of the LCF architecture>
 <either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>
 
-Implementation details to put on the web.
+<Implementation details to put on the web.>
 <simple cgi - worth mentioning?>

### 2012-02-02 paper.tex
minor changes to abstract and intro
based on my memory of mike's comments

@@
 \newcommand{\secref}[1]{Section~\ref{sec:#1}}
+\newcommand{\ie}{i.e.\xspace}
 
@@
 \begin{abstract}
-Tactics, functions that compute low-level inferences to achieve a high-level reasoning goal, are usually developed within and tied to a particular proof assistant, which leads to code duplication across different systems and inhibits reuse.
-As formal methods are applied to larger problems, projects spanning multiple proof assistants benefit more from interoperability and access to the union of all implemented tactics.
+Tactics, \ie functions that compute low-level inferences to achieve a high-level reasoning goal, are usually developed within and tied to a particular proof assistant, which leads to code duplication across different systems and inhibits reuse.
+Projects spanning multiple proof assistants would benefit from interoperability and from access to the union of all implemented tactics.
 Using the OpenTheory proof exchange format, we show how to turn a tactic implemented for one system into one available to many via the web.
@@
 \section{Introduction}
-There are many systems for interactively developing machine-checked formal theories, including HOL4~\cite{slind08brief}, HOL Light~\cite{DBLP:conf/tphol/Harrison09a}, ProofPower~\cite{ProofPower}, Isabelle/HOL~\cite{wenzel08isabelle}, and Coq~\cite{bertot08short}.
-%All but the last are based on higher-order logic~\cite{church40formulation,gordon94hol} and follow the LCF tradition~\cite{gordon79lcf} wherein theorems are represented by an abstract datatype all of whose constructors, representing inference rules, are contained in a small, trusted kernel.
-%In LCF-style provers, powerful proof procedures are kept trustworthy by, ultimately, only calling kernel functions to create theorems.
-The logic implemented by these systems (except Coq) is essentially the same, but the collections of libraries and tactics built atop the logical kernels differ.
+There are many similar LFC-style systems for interactively developing machine-checked formal theories, including HOL4~\cite{slind08brief}, HOL Light~\cite{DBLP:conf/tphol/Harrison09a}, ProofPower~\cite{ProofPower}, and Isabelle/HOL~\cite{wenzel08isabelle}.
+The logic implemented by these systems is essentially the same, but the collections of libraries and tactics built atop the logical kernels differ.
 Where similar functionality exists in two systems it is usually the result of duplicate effort.
@@
 \item
-A standard format (\OpenTheory article files) for goals and proofs, and
+a standard format (\OpenTheory article files) for goals and proofs, and
 \item
-An extensible way to fix the meaning of constants and type definitions.
+an extensible way to fix the meaning of constants and type definitions.
 \end{enumerate}
@@
 \item
-A method (\secref{implementation}) for making existing libraries standalone, using the \OpenTheory article format, including working examples online; and,
+a method (\secref{implementation}) for making existing libraries standalone, using the \OpenTheory article format, including working examples online; and,
 \item
-Preliminary performance data (\secref{performance}) comparing standalone tactics against direct implementations.
+preliminary performance data (\secref{performance}) comparing standalone tactics against direct implementations.
 \end{itemize}

### 2012-02-03 paper.tex
tidying sec 2.1, and start adding example

@@
 The central concept in \OpenTheory is that of a \emph{theory package}, $\Gamma\rhd\Delta$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
-The basic unit\footnote{composite theory packages are also possible} for constructing theory packages is an \emph{article file}, which consists of instructions for a virtual machine whose operations include primitive inference rules of higher-order logic and basic term-manipulation.
-
-<more concrete examples? definition of tactics? viewing tactic as derived inference rule? intro to LCF kernel?>
+The basic unit\footnote{Composite theory packages are also possible.} for constructing theory packages is an \emph{article file}, which consists of instructions for a virtual machine whose operations include primitive inference rules of higher-order logic and basic term-manipulation.
 
@@
 More generally, by taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
-
-Furthermore, we can represent a higher-order logic term $t$ by the theory $\{\vdash\mathtt{K\;F\;}t\}\rhd\{\vdash\mathtt{K\;F\;}t\}$, where $\mathtt{K}:\alpha\to\beta\to\alpha$ is the combinator defined by $\mathtt{K}\;x\;y=x$, and $\mathtt{F}$ is falsity (any boolean would do).
-The article format ensures $t$ is only written once in representing this theory.
-This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.
-
+Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem.
+\footnote{For example, we could represent $t$ by the theory $\{\vdash\mathtt{K\;F\;}t\}\rhd\{\vdash\mathtt{K\;F\;}t\}$, where $\mathtt{K}:\alpha\to\beta\to\alpha$ is the combinator defined by $\mathtt{K}\;x\;y=x$, and $\mathtt{F}$ is falsity (any boolean would do).
+This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.}
 Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
 (Additional inputs and outputs can always be communicated separately.)
+
+As an example, consider the trivial task of proving $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
+Such a task might be given to a standalone tactic for beta-reducing terms.\footnote{In reality, such low-level functionality may not outweigh the overhead of making a tactic standalone.
+See <Sec 4??> for more realistic examples of tactics albeit in less detail.}
+
+<more concrete examples? definition of tactics? viewing tactic as derived inference rule? intro to LCF kernel?>
 

### 2012-02-06 paper.tex
fix typo

@@
 
-Besides duplicate initial effort, costs of maintenance can be multiplied over different provers, and improvements in functionality can become restricted to single prover unnecessarily.
+Besides duplicate initial effort, costs of maintenance can be multiplied over different provers, and improvements in functionality can become restricted to a single prover unnecessarily.
 For instance, Hurd's first-order prover, Metis~\cite{DBLP:conf/cade/Hurd02}, is integrated in at least three HOL-based provers, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.

### 2012-02-06 paper.tex
bit of work on sec 2
concrete, possibly running example, which might lead to a focus on
requirements of systems to be able to make use of standalone tactics

@@
 \subsection{\OpenTheory for Tactic Communication}
+As an example, consider the trivial task of proving $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
+The user of a proof assistant needs to prove this theorem, and decides to pass it off to a standalone tactic for beta-reducing terms.\footnote{In reality, such low-level functionality may not outweigh the overhead of making a tactic stand alone.
+See \secref{performance} for more realistic examples of tactics (in less detail).}
+The input for the tactic is the term $(\lambda{x}.\;x)\;\mathtt{T}$, and the output is a proof of the theorem $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
+Standalone tactics use \OpenTheory article files for communicating both terms and theorems.
+The proof assistant translates the term from its local internal format to \OpenTheory article format, and sends the resulting article over the web to the tactic.
+The tactic returns another article, encoding a proof, which the proof assistant replays to create the desired theorem.
+
+<Example illustrates requirements of the proof assistant: be able to read proofs from articles, be able to write terms as articles, and some form of communication with external program or website>
+
 The central concept in \OpenTheory is that of a \emph{theory package}, $\Gamma\rhd\Delta$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
@@
 Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
-(Additional inputs and outputs can always be communicated separately.)
-
-As an example, consider the trivial task of proving $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
-Such a task might be given to a standalone tactic for beta-reducing terms.\footnote{In reality, such low-level functionality may not outweigh the overhead of making a tactic standalone.
-See <Sec 4??> for more realistic examples of tactics albeit in less detail.}
-
-<more concrete examples? definition of tactics? viewing tactic as derived inference rule? intro to LCF kernel?>
+Additional data can be communicated separately; for example, the beta-reducing tactic might take a parameter specifying the maximum number of reductions to perform.
 

### 2012-02-07 paper.tex
more on 2.1

@@
 
-<Example illustrates requirements of the proof assistant: be able to read proofs from articles, be able to write terms as articles, and some form of communication with external program or website>
+The example illustrates the key requirements of the proof assistant: the ability to
+\begin{enumerate}
+\item \label{req:artr} replay proofs by reading \OpenTheory articles,
+\item \label{req:artw} write terms as \OpenTheory articles, and,
+\item \label{req:comm} communicate, for example over the web, with external programs.
+\end{enumerate}
 
+Items~\ref{req:artr} and \ref{req:artw} require implementation work for interpreting and constructing articles.
 The central concept in \OpenTheory is that of a \emph{theory package}, $\Gamma\rhd\Delta$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
-The basic unit\footnote{Composite theory packages are also possible.} for constructing theory packages is an \emph{article file}, which consists of instructions for a virtual machine whose operations include primitive inference rules of higher-order logic and basic term-manipulation.
+An \emph{article file} is the basic unit\footnote{Composite theory packages are also possible.} for constructing theory packages.
+It consists of instructions for a virtual machine whose operations include primitive inference rules of higher-order logic and basic term-manipulation.
+
+The instructions allowed in an article file represent the \OpenTheory \emph{logical kernel}.
+To read an article, a proof assistant performs the inferences and other instructions listed in the file.
+The \OpenTheory kernel is very close to HOL Light's kernel.
+It should be easy to read articles in any proof assistant that manipulates higher-order logic.
 
@@
 Additional data can be communicated separately; for example, the beta-reducing tactic might take a parameter specifying the maximum number of reductions to perform.
+
+<why is it good for communicating terms and theorems? fixes ontology...>
 

### 2012-02-07 paper.tex
bit more paper hacking

@@
 An article file represents a theory $\Gamma\rhd\Delta$.
-As a special case, we can represent the proof of a single goal $g$ by taking $\Delta=\{\vdash g\}$.
-More generally, by taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
+By taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
 Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem.
@@
 
-<why is it good for communicating terms and theorems? fixes ontology...>
+The main advantage to using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems. <Example?>
+<Requirements~\ref{req:artr} and \ref{req:artw} above carry the additional benefit of giving a proof assistant access to traditional \OpenTheory packages (\ie not just the output of standalone tactics).>
 
@@
 We have experimented with the second approach, taking tactics from HOL4 and from HOL Light.
-<The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.>
+The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.
 The key requirements of the proof assistant are:
@@
 \item
-An \OpenTheory article file reader and writer,
+\label{req:io} an \OpenTheory article file reader and writer,
 \item
-A facility for recording proofs and converting them to use the \OpenTheory primitive inferences, and,
+\label{req:log} a facility for recording proofs and converting them to the \OpenTheory logical kernel, and,
 \item
-A facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
+\label{req:ckpt} a facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
 \end{enumerate}
-<first two items are the most work, but have the additional benefit of allowing import/export of theories in opentheory repositories>
+<first two items are the most work, but have the additional benefit of allowing import/export of theories in opentheory repositories, as mentioned above.
+so: to use standalone tactics you also get to use opentheory packages, but if you can make yourself into a standalone tactic you can also export opentheory packages.>
 <in HOL4 and HOL Light, item 2 is accomplished by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
-this step is made easy because of the LCF architecture>
+this step is made easy by LCF architecture: could be very difficult in systems that don't already generate proofs somehow, or with significantly different kernel (cf isabelle)>
 <either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>
 
-<Implementation details to put on the web.>
-<simple cgi - worth mentioning?>
+<Implementation details to put on the web: simple cgi>
 
@@
 \section{Conclusions}
-Additional (non-proof) work tactics might also do needs to be represented on another channel.
 

### 2012-02-08 paper.tex
change the example in 2.1 to something less trivial

@@
 \subsection{\OpenTheory for Tactic Communication}
-As an example, consider the trivial task of proving $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
-The user of a proof assistant needs to prove this theorem, and decides to pass it off to a standalone tactic for beta-reducing terms.\footnote{In reality, such low-level functionality may not outweigh the overhead of making a tactic stand alone.
-See \secref{performance} for more realistic examples of tactics (in less detail).}
-The input for the tactic is the term $(\lambda{x}.\;x)\;\mathtt{T}$, and the output is a proof of the theorem $\vdash(\lambda{x}.\;x)\;\mathtt{T}=\mathtt{T}$.
+As an example, consider proving $\vdash20+(12\times x)<(x\times 8)-13$.
+The user of a proof assistant needs to prove this theorem, and decides to pass it off to a standalone tactic for linear arithmetic.
+The input for the tactic is the term $20+(12\times x)<(x\times 8)-13$, and the output is a proof of the theorem.
 Standalone tactics use \OpenTheory article files for communicating both terms and theorems.
@@
 The \OpenTheory kernel is very close to HOL Light's kernel.
-It should be easy to read articles in any proof assistant that manipulates higher-order logic.
+It should be easy to read articles in any system that can prove theorems of higher-order logic.
 
@@
 By taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
+In our example above, we have $\Delta=\{\vdash20+(12\times x)<(x\times 8)-13\}$ and might have $\Gamma\ni\forall{m\,n}.\;m\times n = n\times m$ amongst other assumptions.
 Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem.
@@
 Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
-Additional data can be communicated separately; for example, the beta-reducing tactic might take a parameter specifying the maximum number of reductions to perform.
+Additional (non-logical) data can always be communicated separately.
 
-The main advantage to using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems. <Example?>
-<Requirements~\ref{req:artr} and \ref{req:artw} above carry the additional benefit of giving a proof assistant access to traditional \OpenTheory packages (\ie not just the output of standalone tactics).>
+The main advantage to using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems.
+For example, the term $20+(12\times x)<(x\times 8)-13$ might involve applications of constants \texttt{Number.Numeral.zero} and \texttt{Number.Numeral.bit1}, encoding the numbers in binary format.
+The names and namespaces of these constants are indicated in the \OpenTheory standard library, and proof assistants can maintain translations to and from their local names.
+A system using a different encoding for numbers (say unary, with \texttt{Number.Natural.suc}) could use additional standalone tactics to translate between encodings.
+<...?>
+
+Requirements~\ref{req:artr} and \ref{req:artw} above carry the additional benefit of giving a proof assistant access to traditional \OpenTheory packages (\ie not just the output of standalone tactics).
 

### 2012-02-08 paper.tex
work on sec 2.2

@@
 \newcommand{\ie}{i.e.\xspace}
+\newcommand{\eg}{e.g.\xspace}
 
@@
 \end{enumerate}
-<first two items are the most work, but have the additional benefit of allowing import/export of theories in opentheory repositories, as mentioned above.
-so: to use standalone tactics you also get to use opentheory packages, but if you can make yourself into a standalone tactic you can also export opentheory packages.>
-<in HOL4 and HOL Light, item 2 is accomplished by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
-this step is made easy by LCF architecture: could be very difficult in systems that don't already generate proofs somehow, or with significantly different kernel (cf isabelle)>
+
+Just as the requirements of a system to use standalone tactics also enable it to import \OpenTheory packages in general, the requirements to create standalone tactics (specifically \ref{req:io} and \ref{req:log}) also enable a system to create and export \OpenTheory packages.
+
+Requirement~\ref{req:log} above poses the highest barrier: a standalone tactic must record proofs at a level of detail sufficient to create proofs of the same theorems using the \OpenTheory kernel.
+It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules.
+In HOL4 and HOL Light, we met Requirement~\ref{req:log} by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
+Then we express each primitive inference as a derived rule above the \OpenTheory kernel.
+In many cases, there is a direct counterpart (\eg reflexivity of equality is primitive in all three kernels).
+But, for example, HOL4's primitive rule for definition by specification must be emulated by a definition using Hilbert choice in \OpenTheory.
+<could be difficult in system with significantly different kernel, eg isabelle>
+
 <either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>

### 2012-02-09 paper.tex
fleshing out sec 2.2

@@
 But, for example, HOL4's primitive rule for definition by specification must be emulated by a definition using Hilbert choice in \OpenTheory.
-<could be difficult in system with significantly different kernel, eg isabelle>
+Although Isabelle/HOL uses the LCF architecture, its logical kernel is quite different from \OpenTheory; we therefore expect translating Isabelle/HOL proofs to be more difficult than HOL4 proofs.
 
-<either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>
-
-<Implementation details to put on the web: simple cgi>
+To satisfy Requirement~\ref{req:ckpt} in HOL4, we used the ``export'' facility of the PolyML compiler, which creates a standalone executable that runs an ML function.
+(For each tactic, we captured a function that reads an article representing the input term, runs the tactic (recording the proof), and writes an article containing the proof.)
+The situation for HOL Light is more complicated because OCaml does not provide such an ``export'' facility, and HOL Light does not use stored theories.
+We used a general purpose checkpointing facility to capture a HOL Light session at an appropriate point (ready to read an article and run a tactic).
 
@@
 Omega.
+MathServe?
 Natural extension of the integration of automatic tools into interactive systems.
@@
 
+Using \OpenTheory for communication, we can
+
 Highly customised tactics for a particular proof system might suffer performance hits if the factorisation through the OpenTheory kernel cannot be done efficiently.
-By the same token, the article format might admit optimisation by restructuring low-level proofs, and such optimisers could be widely applied.
+<The article format might admit optimisation by restructuring low-level proofs, and such optimisers could be widely applied.>
+<Tactics where proof search dominates proof replay offer the best prospects, since the search can be done elsewhere (in parallel), the overhead is mainly in transporting and replaying proofs, and queries might be cached>
 

### 2012-02-10 paper.tex
scribblings

@@
 
+<mention the need in some cases to filter articles through the opentheory tool?
+maybe do this in the performance section.
+the reasons include: to remove prover-specific (Unwanted.id) tags, and to canonicalise HOL Light's constant names (using the interpret: as: package lines).
+Overall, these details are probably unimportant...
+but affect the performance>
+
 \section{Preliminary Performance Results}
@@
 
+<The tactics currently available, what they do, and where to find them.>
+
 Speed comparisons against using built-in tactics.
 <compare builtin holqbf, holqbf with builtin minisat, and holqbf with external minisat (i.e. chain of two webtactics) on a simple and difficult qbf problem>
+
+<How does (or would we expect) performance to scale with size of terms and proofs involved? Encoding/decoding terms/proofs presumably scales linearly.>
 
@@
 
-Using \OpenTheory for communication, we can
+We have shown how, using \OpenTheory for communication, we can make tactics for higher-order logic reasoning tasks available to multiple proof assistants and independently maintainable.
 
-Highly customised tactics for a particular proof system might suffer performance hits if the factorisation through the OpenTheory kernel cannot be done efficiently.
+<Proof assistants enable the development of formal theories both by providing an environment for managing formalisations, that is, keeping track of the details and maintaning consistency, and also by providing customised tools for specific reasoning tasks.>
+
+<Tactics tailored to a particular proof system would suffer performance hits if the factorisation through the OpenTheory kernel cannot be done efficiently.>
 <The article format might admit optimisation by restructuring low-level proofs, and such optimisers could be widely applied.>
@@
 
-Other examples of functionality that could usefully be made prover-independent include packages for defining recursive datatypes or for solving classes of arithmetic problems.
+<Other examples of functionality that could usefully be made prover-independent include packages for defining recursive datatypes, quotient types, nominal types, or for solving classes of arithmetic problems (decision procedures).
+- worth mentioning?
+Better to only list future work one actually intends to do.
+(E.g.
+isabelle and sledgehammer)>
 

### 2012-02-11 paper.tex
tweaks

@@
 The main advantage to using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems.
-For example, the term $20+(12\times x)<(x\times 8)-13$ might involve applications of constants \texttt{Number.Numeral.zero} and \texttt{Number.Numeral.bit1}, encoding the numbers in binary format.
+For example, the term $20+(12\times x)<(x\times 8)-13$ might involve applications of constants \texttt{Number.Natural.bit0} and \texttt{Number.Natural.bit1}, encoding the numbers in binary format.
 The names and namespaces of these constants are indicated in the \OpenTheory standard library, and proof assistants can maintain translations to and from their local names.
-A system using a different encoding for numbers (say unary, with \texttt{Number.Natural.suc}) could use additional standalone tactics to translate between encodings.
-<...?>
+A system using a different encoding for numbers (say unary, with \texttt{Number.Natural.suc} and \texttt{Number.Natural.zero}) could use additional standalone tactics to translate between encodings.
 
@@
 Requirement~\ref{req:log} above poses the highest barrier: a standalone tactic must record proofs at a level of detail sufficient to create proofs of the same theorems using the \OpenTheory kernel.
-It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules.
-In HOL4 and HOL Light, we met Requirement~\ref{req:log} by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
-Then we express each primitive inference as a derived rule above the \OpenTheory kernel.
+It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules, as is the case for HOL4 and HOL Light.
+
+We met Requirement~\ref{req:log} by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
+Then we expressed each primitive inference as a derived rule above the \OpenTheory kernel.
 In many cases, there is a direct counterpart (\eg reflexivity of equality is primitive in all three kernels).
-But, for example, HOL4's primitive rule for definition by specification must be emulated by a definition using Hilbert choice in \OpenTheory.
-Although Isabelle/HOL uses the LCF architecture, its logical kernel is quite different from \OpenTheory; we therefore expect translating Isabelle/HOL proofs to be more difficult than HOL4 proofs.
+But, for example, HOL4's primitive rule for definition by specification must be emulated in \OpenTheory, for example by using Hilbert choice.
+Although Isabelle/HOL uses the LCF architecture, its logical kernel is quite different from \OpenTheory; we therefore expect translating Isabelle/HOL proofs to be more difficult than HOL4 and HOL Light proofs.
 

### 2012-02-11 paper.tex
some sentences for sec 4

@@
 
-System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}.
-<Comparison to THF?>
+An impressive example of providing reasoning infrastructure over the web is the System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}.
+Here, multiple automatic theorem provers can be called on a problem (\eg satisfiability of a formula, or whether a goal follows from assumptions) expressed in the standard TPTP format.
+The usual scenario is a decision problem in first-order logic, but TPTP includes support for higher-order terms (THF format) and for expressing proofs in a standard language (TSTP).
+
+<idea of separate reasoning tools communicating and enhancing a proof development environment>
 Evidential tool bus~\cite{DBLP:conf/icfem/Rushby05,fk:etb}.
@@
 MathServe?
-Natural extension of the integration of automatic tools into interactive systems.
+<This is a natural extension of the integration of automatic tools into interactive systems.>
 

### 2012-02-11 paper.tex
bit more sec 4

@@
 
-<idea of separate reasoning tools communicating and enhancing a proof development environment>
-Evidential tool bus~\cite{DBLP:conf/icfem/Rushby05,fk:etb}.
-Omega.
-MathServe?
-<This is a natural extension of the integration of automatic tools into interactive systems.>
+The idea of separate reasoning tools communicating to enhance a proof development environment is also being pursued in the Evidential Tool Bus~\cite{DBLP:conf/icfem/Rushby05,fk:etb} and the MathServe System~\cite{DBLP:conf/cade/ZimmerA06}.
+This idea is a natural extension of the integration of automatic tools with interactive theorem provers.
 

### 2012-02-12 paper.tex
use cleaner term-encoding example in the footnote, avoid mentioning K

@@
 In our example above, we have $\Delta=\{\vdash20+(12\times x)<(x\times 8)-13\}$ and might have $\Gamma\ni\forall{m\,n}.\;m\times n = n\times m$ amongst other assumptions.
-Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem.
-\footnote{For example, we could represent $t$ by the theory $\{\vdash\mathtt{K\;F\;}t\}\rhd\{\vdash\mathtt{K\;F\;}t\}$, where $\mathtt{K}:\alpha\to\beta\to\alpha$ is the combinator defined by $\mathtt{K}\;x\;y=x$, and $\mathtt{F}$ is falsity (any boolean would do).
+Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem.\footnote{For example, we could represent $t$ by the theory $\emptyset\rhd\{\vdash{t=t}\}$.
+(The article format allows for multiple occurrences of $t$ to be shared.)
 This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.}

### 2012-02-13 paper.tex
Joe's pass

@@
 \newcommand{\OpenTheory}{OpenTheory\xspace}
+\newcommand{\theory}[2]{{#1}\,\triangleright\,{#2}}
 \newcommand{\secref}[1]{Section~\ref{sec:#1}}
@@
 
-\institute{\email{Ramana.Kumar@cl.cam.ac.uk}\and\email{joe@galois.com}}
+\institute{University of Cambridge, \email{Ramana.Kumar@cl.cam.ac.uk}\and Galois, Inc., \email{joe@gilith.com}}
 
@@
 \begin{abstract}
-Tactics, \ie functions that compute low-level inferences to achieve a high-level reasoning goal, are usually developed within and tied to a particular proof assistant, which leads to code duplication across different systems and inhibits reuse.
-Projects spanning multiple proof assistants would benefit from interoperability and from access to the union of all implemented tactics.
-Using the OpenTheory proof exchange format, we show how to turn a tactic implemented for one system into one available to many via the web.
-This enables, for example, LCF-style proof reconstruction efforts for fast automatic provers to be shared by users of multiple proof assistants and removes the need for each user to install the automatic prover.
+Proof tools in interactive theorem provers are usually developed within and tied to a specific system, which leads to a duplication of effort to make the functionality available in different systems.
+Many verification projects would benefit from access to the union of all proof tools implemented across interactive theorem provers.
+Using \OpenTheory as a language for communicating between systems, we show how to turn an existing proof tool implemented for one system into a standalone tactic available to many systems via the web.
+This enables, for example, LCF-style proof reconstruction efforts for integrated external proof tools to be shared by users of different interactive theorem provers and removes the need for each user to install the external proof tool.
 \end{abstract}
@@
 \section{Introduction}
-There are many similar LFC-style systems for interactively developing machine-checked formal theories, including HOL4~\cite{slind08brief}, HOL Light~\cite{DBLP:conf/tphol/Harrison09a}, ProofPower~\cite{ProofPower}, and Isabelle/HOL~\cite{wenzel08isabelle}.
-The logic implemented by these systems is essentially the same, but the collections of libraries and tactics built atop the logical kernels differ.
-Where similar functionality exists in two systems it is usually the result of duplicate effort.
 
-It makes sense to speak of similar functionality in different systems not just because they use roughly the same logic, but also because there is a shared base of concepts: booleans, inductive datatypes, recursive functions, natural numbers, lists, sets, etc.
-The \OpenTheory standard library~\cite{DBLP:conf/nfm/Hurd11} formalises this shared base, using \OpenTheory~\cite{hurd2009}, a format designed for storing and sharing higher-order logic theories.
-We use \OpenTheory to communicate with tactics (sending goals and receiving proofs), and thereby obtain
-\begin{enumerate}
-\item
-a standard format (\OpenTheory article files) for goals and proofs, and
-\item
-an extensible way to fix the meaning of constants and type definitions.
-\end{enumerate}
+There are many LCF-style systems for interactively developing machine-checked formal theories, including HOL4~\cite{slind08brief}, HOL Light~\cite{DBLP:conf/tphol/Harrison09a}, ProofPower~\cite{ProofPower}, and Isabelle/HOL~\cite{wenzel08isabelle}.
+The logic implemented by these systems is essentially the same, but the collections of theory libraries and proof tools built on top of the logical kernels differ.
+Where similar proof tools exist in multiple systems it is usually the result of duplicated effort.
 
@@
 Besides duplicate initial effort, costs of maintenance can be multiplied over different provers, and improvements in functionality can become restricted to a single prover unnecessarily.
-For instance, Hurd's first-order prover, Metis~\cite{DBLP:conf/cade/Hurd02}, is integrated in at least three HOL-based provers, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
-Another example is a package for defining recursive functions, TFL~\cite{slindphd}, originally implemented for both Isabelle/HOL and HOL4, but recently superceded in Isabelle/HOL only by Krauss's function definition package~\cite{DBLP:journals/jar/Krauss10}.
-The improvements of Krauss's method over TFL ought to be applicable to other HOL-based provers, but a direct re-implementation would require substantial effort.
+For instance, the Metis first-order prover~\cite{hurd2003d} is integrated in at least three HOL-based provers, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
+Another example is Slind's TFL package for defining recursive functions~\cite{slindphd}, originally implemented for both Isabelle/HOL and HOL4, and recently superceded in Isabelle/HOL by Krauss's function definition package~\cite{DBLP:journals/jar/Krauss10}.
+The improvements of Krauss's method over TFL ought to be applicable to other HOL-based provers, but a direct reimplementation would require substantial effort.
 
-We contend that tactics or libraries for proof assistants need only be written and maintained in one place rather than once per system. 
-An added advantage when the tactic is an integration of an external tool is that a user of the proof assistant need not also have installed the external tool, because a standalone tactic can be accessed from a web server where the tool is installed.
+It makes sense to speak of similar proof tools in different interactive theorem provers not just because they implement essentially the same logic, but also because there is a shared base of concepts: booleans, inductive datatypes, recursive functions, natural numbers, lists, sets, etc.
+The \OpenTheory standard library~\cite{hurd2011} formalises this shared base as a collection of theory packages containing proofs written in the simple \emph{article} format designed for storing and sharing higher order logic theories~\cite{hurd2009}.
+We use \OpenTheory as a language for users of interactive theorem provers to communicate with proof tools on a remote server, and thereby obtain the following two benefits:
+\begin{enumerate}
+\item\textbf{Proof Articles:} A standard format to encode the goals that will be sent to the proof tools and the proofs that will be received.
+
+\item\textbf{Standard Library:} An extensible way to fix the meaning of constants and type definitions between systems.
+\end{enumerate}
+
+We contend that proof tools for interactive theorem provers need only be written and maintained in one place rather than once per system, using \emph{standalone tactics} that are available on the internet and use \OpenTheory to communicate.
+An added advantage when the tactic is an integration of an external tool is that a user of the interactive theorem prover need not also have installed the external tool, because the standalone tactic can be accessed from a web server where the tool is installed.
 
 The contributions of this rough diamond are:
-\begin{itemize}
-\item
-a method (\secref{implementation}) for making existing libraries standalone, using the \OpenTheory article format, including working examples online; and,
-\item
-preliminary performance data (\secref{performance}) comparing standalone tactics against direct implementations.
-\end{itemize}
+\begin{enumerate}
+\item A general method for turning existing proof tools implemented in interactive theorem provers into standalone tactics~(\secref{implementation}).
 
-\section{Making Existing Tactics into Web Services}
+\item Preliminary results profiling the performance of working examples of standalone tactics~(\secref{performance}).
+\end{enumerate}
+
+\section{Lifting Proof Tools into the Cloud}
 \label{sec:implementation}
@@
 \subsection{\OpenTheory for Tactic Communication}
-As an example, consider proving $\vdash20+(12\times x)<(x\times 8)-13$.
-The user of a proof assistant needs to prove this theorem, and decides to pass it off to a standalone tactic for linear arithmetic.
-The input for the tactic is the term $20+(12\times x)<(x\times 8)-13$, and the output is a proof of the theorem.
-Standalone tactics use \OpenTheory article files for communicating both terms and theorems.
-The proof assistant translates the term from its local internal format to \OpenTheory article format, and sends the resulting article over the web to the tactic.
-The tactic returns another article, encoding a proof, which the proof assistant replays to create the desired theorem.
 
-The example illustrates the key requirements of the proof assistant: the ability to
+As an example, suppose the user of an interactive theorem prover faced with the goal
+\[
+\forall n.\ 8\le n\Rightarrow\exists s,t.\ n = 3s + 5t
+\]
+decides to pass it off to a standalone tactic for linear arithmetic.\footnote{This is a representation of the problem asking for a proof that \pounds 3 and \pounds 5 coins can be used to make \pounds $n$ for any natural number $n$ that is at least 8.}
+
+The input for the standalone tactic is the goal, and the output is a proof of the theorem.
+Standalone tactics use the \OpenTheory article format for communicating both goal terms and proofs of theorems.
+The interactive theorem prover serializes the goal term from its local internal format to an article file, and sends the article over the web to the standalone tactic.
+If successful, the standalone tactic sends back another article file encoding a proof of the goal, which the interactive theorem prover replays through its logical kernel to create the desired theorem.
+
+This example illustrates the key requirements for an interactive theorem prover to use standalone tactics:
 \begin{enumerate}
-\item \label{req:artr} replay proofs by reading \OpenTheory articles,
-\item \label{req:artw} write terms as \OpenTheory articles, and,
-\item \label{req:comm} communicate, for example over the web, with external programs.
+\item\label{req:artr} Ability to replay proofs by reading \OpenTheory articles.
+
+\item\label{req:artw} Ability to write terms as \OpenTheory articles.
+
+\item\label{req:comm} Ability to communicate over the web with external programs.
 \end{enumerate}
 
-Items~\ref{req:artr} and \ref{req:artw} require implementation work for interpreting and constructing articles.
-The central concept in \OpenTheory is that of a \emph{theory package}, $\Gamma\rhd\Delta$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
-An \emph{article file} is the basic unit\footnote{Composite theory packages are also possible.} for constructing theory packages.
-It consists of instructions for a virtual machine whose operations include primitive inference rules of higher-order logic and basic term-manipulation.
+Requirements~\ref{req:artr} and \ref{req:artw} can be satisfied for an interactive theorem prover by implementing an \OpenTheory interface that can interpret and construct articles.
+The central concept in \OpenTheory is that of a \emph{theory package}, $\theory{\Gamma}{\Delta}$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
+An article is a concrete representation of a theory package, consisting of instructions for a virtual machine whose operations include construction of types, terms, and the primitive inference rules of higher order logic.
+To read an article, an interactive theorem prover performs the primitive inferences and other instructions listed in the file.
+The \OpenTheory logical kernel is based on HOL Light's logical kernel, and the instructions are chosen to make it easy to read articles into any system that can prove theorems of higher order logic.
 
-The instructions allowed in an article file represent the \OpenTheory \emph{logical kernel}.
-To read an article, a proof assistant performs the inferences and other instructions listed in the file.
-The \OpenTheory kernel is very close to HOL Light's kernel.
-It should be easy to read articles in any system that can prove theorems of higher-order logic.
+An article file represents a theory $\theory{\Gamma}{\Delta}$.
+By taking $\Delta$ to be the set of theorems proved by a proof tool and $\Gamma$ to be the set of theorems used by the proof tool, we can view the result of executing a proof tool as a logical theory.
+In our example above of using a linear arithmetic standalone tactic implementing on the given goal, this theory might be
+\[
+\theory{\left\{\begin{array}{l}
+\vdash\forall n.\ n + 0 = n \\
+\vdash\forall m,n.\ m + n = n + m \\
+\cdots
+\end{array}\right\}}
+{\bigl\{\vdash\forall n.\ 8\le n\Rightarrow\exists s,t.\ n = 3s + 5t\bigr\}}
+\]
+where the assumptions consist of a collection of standard arithmetic facts.
 
-An article file represents a theory $\Gamma\rhd\Delta$.
-By taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
-In our example above, we have $\Delta=\{\vdash20+(12\times x)<(x\times 8)-13\}$ and might have $\Gamma\ni\forall{m\,n}.\;m\times n = n\times m$ amongst other assumptions.
-Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem.\footnote{For example, we could represent $t$ by the theory $\emptyset\rhd\{\vdash{t=t}\}$.
-(The article format allows for multiple occurrences of $t$ to be shared.)
-This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.}
-Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
-Additional (non-logical) data can always be communicated separately.
+% Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem.\footnote{For example, we could represent $t$ by the theory $\emptyset\rhd\{\vdash{t=t}\}$.
+% (The article format allows for multiple occurrences of $t$ to be shared.)
+% This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.}
+% Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
+% Additional (non-logical) data can always be communicated separately.
 
 The main advantage to using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems.
-For example, the term $20+(12\times x)<(x\times 8)-13$ might involve applications of constants \texttt{Number.Natural.bit0} and \texttt{Number.Natural.bit1}, encoding the numbers in binary format.
-The names and namespaces of these constants are indicated in the \OpenTheory standard library, and proof assistants can maintain translations to and from their local names.
+For example, the numerals 3, 5 and 8 in the example goal term can be encoded in binary using the standard constants \texttt{Number.Natural.bit0} and \texttt{Number.Natural.bit1}.
+The full names and properties of these constants are indicated in the \OpenTheory standard library, and interactive theorem provers can maintain translations to and from their local names and theorems.
 A system using a different encoding for numbers (say unary, with \texttt{Number.Natural.suc} and \texttt{Number.Natural.zero}) could use additional standalone tactics to translate between encodings.
 
-Requirements~\ref{req:artr} and \ref{req:artw} above carry the additional benefit of giving a proof assistant access to traditional \OpenTheory packages (\ie not just the output of standalone tactics).
+Implementing an \OpenTheory interface to satisfy Requirements~\ref{req:artr} and~\ref{req:artw} above carries the additional benefit of giving the interactive theorem prover access to all theories stored as \OpenTheory packages, not just those that are the output of standalone tactics.
 
-\subsection{Isolating tactics from their home systems}
-There are two approaches to obtaining a standalone tactic: either write one directly, or extract an existing tactic from a proof assistant. 
-We have experimented with the second approach, taking tactics from HOL4 and from HOL Light.
+\subsection{Extracting Tactics from Interactive Theorem Provers}
+
+There are two approaches to obtaining a standalone tactic: either write one directly; or extract an existing tactic from an interactive theorem prover.
+We have experimented with the second approach, extracting tactics from HOL4 and from HOL Light.
 The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.
-The key requirements of the proof assistant are:
+For tactic extraction to succeed, the key requirements of the interactive theorem prover are:
 \begin{enumerate}
-\item
-\label{req:io} an \OpenTheory article file reader and writer,
-\item
-\label{req:log} a facility for recording proofs and converting them to the \OpenTheory logical kernel, and,
-\item
-\label{req:ckpt} a facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
+\item\label{req:io} An \OpenTheory article file reader and writer.
+
+\item\label{req:log} A facility for recording proofs and converting them to the \OpenTheory logical kernel.
+
+\item\label{req:ckpt} A facility for check-pointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
 \end{enumerate}
 
-Just as the requirements of a system to use standalone tactics also enable it to import \OpenTheory packages in general, the requirements to create standalone tactics (specifically \ref{req:io} and \ref{req:log}) also enable a system to create and export \OpenTheory packages.
+Just as the requirements of an interactive theorem prover to use standalone tactics also enable it to import \OpenTheory packages in general, the requirements to create standalone tactics (specifically Requirements~\ref{req:io} and~\ref{req:log}) also enable a system to create and export \OpenTheory packages.
 
@@
 
-To satisfy Requirement~\ref{req:ckpt} in HOL4, we used the ``export'' facility of the PolyML compiler, which creates a standalone executable that runs an ML function.
-(For each tactic, we captured a function that reads an article representing the input term, runs the tactic (recording the proof), and writes an article containing the proof.)
-The situation for HOL Light is more complicated because OCaml does not provide such an ``export'' facility, and HOL Light does not use stored theories.
-We used a general purpose checkpointing facility to capture a HOL Light session at an appropriate point (ready to read an article and run a tactic).
+To satisfy Requirement~\ref{req:ckpt} in HOL4, we used the `export' facility of the PolyML compiler, which creates a standalone executable that runs an ML function.
+For each tactic, we captured a function that reads an article representing the input term, runs the tactic (recording the proof), and writes an article containing the proof.
+The situation for HOL Light is more complicated because OCaml does not provide such an `export' facility, and HOL Light does not use stored theories.
+We used a general-purpose checkpointing facility to capture a HOL Light process at the point where it is ready to read an article and run a tactic.
 
-<mention the need in some cases to filter articles through the opentheory tool?
+% <mention the need in some cases to filter articles through the opentheory tool?
 maybe do this in the performance section.
@@
 
-<The tactics currently available, what they do, and where to find them.>
+\begin{table}
+\begin{center}
+\begin{tabular}{||@{~}l@{~}||@{}l@{~}r@{~}|@{~}r@{~}l@{}||@{~}r@{~}||@{}l@{~}r@{~}|@{~}r@{~}l@{}||@{~}r@{~}||}
+\hline
+\textbf{Problem} & & \multicolumn{2}{c}{\textbf{Goal}} & & \textbf{Web} & & \multicolumn{2}{c}{\textbf{Proof}} & & \textbf{Total} \\
+\cline{2-4} \cline{7-9}
+\textbf{Name} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} \\
+ & & \textbf{(b)} & \textbf{(s)} & & \textbf{(s)} & & \textbf{(b)} & \textbf{(s)} & & \textbf{(s)} \\
+\hline
+QBF${}_1$ & & 927 & 0.001 & & 1.064 & & 10,991 & 0.022 & & 1.088 \\
+QBF${}_2$ & & 1,474 & 0.001 & & 1.892 & & 79,944 & 0.139 & & 2.034 \\
+QBF${}_3$ & & 1,546 & 0.001 & & 1.821 & & 91,639 & 0.172 & & 1.996 \\
+SKICo${}_1$ & & 927 & 0.001 & & 1.212 & & 20,047 & 0.041 & & 1.255 \\
+SKICo${}_2$ & & 1,474 & 0.002 & & 1.557 & & 52,249 & 0.113 & & 1.673 \\
+SKICo${}_3$ & & 1,546 & 0.002 & & 1.716 & & 60,642 & 0.125 & & 1.844 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Performance profiling for the test standalone tactics.}
+\label{tab:performance}
+\end{table}
 
-Speed comparisons against using built-in tactics.
-<compare builtin holqbf, holqbf with builtin minisat, and holqbf with external minisat (i.e. chain of two webtactics) on a simple and difficult qbf problem>
+\textbf{RAMANA: perhaps add a column for the time taken to execute the tactic locally and talk about it briefly in the execution time paragraph below.}
 
-<How does (or would we expect) performance to scale with size of terms and proofs involved? Encoding/decoding terms/proofs presumably scales linearly.>
+Table~\ref{tab:performance} shows preliminary performance results executing the QBF and SKICo standalone tactics on the following three test goals over boolean variables:
+\begin{enumerate}
+\item $\forall x.\ x \lor\lnot x$
 
-Amount of code and time required, compared to porting tactics at the implementation level.
+\item $\exists p.\ (\forall q.\ p \lor\lnot q) \land\forall q.\ \exists r.\ r$
+
+\item $\exists x.\ \forall y.\ \exists z.\ (\lnot x \lor\lnot y) \land (\lnot z\lor\lnot y)$
+\end{enumerate}
+For each invocation of a standalone tactic on a test goal, Table~\ref{tab:performance} profiles the time and space requirements of the three phases of execution: encoding the goal as an article; communicating and executing the standalone tactic via the web; and replaying the proof article received.
+
+Table~\ref{tab:performance} shows the size of the articles for the test goals and the resulting proofs is comparable to the typical size of web requests and the resulting pages, so we can be confident that we are within the normal operating range of the web tools we use (curl on the client and CGI scripts on the server).
+For standalone tactics generating large proofs we may wish to compress them using the \texttt{gzip} tool before sending them to the client---previous experiments showed a 90\% compression ratio is typical on articles~\cite{hurd2009}.
+
+As for the execution time, Table~\ref{tab:performance} shows the bulk of the time is spent communicating and executing the standalone tactic via the web, so there is not much profit to be gained in optimizing the client.
+Using the \texttt{traceroute} tool we see a 0.173s network delay between the test client in Portland, OR, USA and the test server in Cambridge, UK, which accounts for at least 0.346s of delay.
+The overall time is in the 1--2s range, which is very slow for workaday tactics but may well be tolerated by the user to gain access to the functionality of a proof tool on another system.
+
+% Speed comparisons against using built-in tactics.
+% <compare builtin holqbf, holqbf with builtin minisat, and holqbf with external minisat (i.e. chain of two webtactics) on a simple and difficult qbf problem>
+
+% <How does (or would we expect) performance to scale with size of terms and proofs involved? Encoding/decoding terms/proofs presumably scales linearly.>
+
+% Amount of code and time required, compared to porting tactics at the implementation level.
 

### 2012-02-13 paper.tex
pick 1 reference for sledgehammer (from 3)

@@
 Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems (HOL4, HOL Light, and Isabelle/HOL); each integration requires a separate implementation.
-Sledgehammer~\cite{DBLP:journals/japll/MengP09,Paulson_threeyears,DBLP:conf/frocos/BlanchetteBN11} is only available for Isabelle/HOL, but would be useful if available for other provers.
+Sledgehammer~\cite{DBLP:conf/frocos/BlanchetteBN11} is only available for Isabelle/HOL, but would be useful if available for other provers.
 

### 2012-02-13 paper.tex
tweak a couple of sentences in intro

@@
 
-We contend that proof tools for interactive theorem provers need only be written and maintained in one place rather than once per system, using \emph{standalone tactics} that are available on the internet and use \OpenTheory to communicate.
-An added advantage when the tactic is an integration of an external tool is that a user of the interactive theorem prover need not also have installed the external tool, because the standalone tactic can be accessed from a web server where the tool is installed.
+We contend that proof tools for interactive theorem provers need only be written and maintained in one place rather than once per system, using \emph{standalone tactics} that are available online and communicate using \OpenTheory.
+An added advantage when the tactic is an integration of an external tool is that a user of the interactive theorem prover need not install the external tool: it only needs to be available on the server hosting the standalone tactic.
 

### 2012-02-13 paper.tex
minor tweaks

@@
 
-As an example, suppose the user of an interactive theorem prover faced with the goal
+As an example, suppose the user of an interactive theorem prover is faced with the goal
 \[
@@
 The central concept in \OpenTheory is that of a \emph{theory package}, $\theory{\Gamma}{\Delta}$, which proves that the set of theorems $\Delta$ logically derive from the set of assumptions $\Gamma$.
-An article is a concrete representation of a theory package, consisting of instructions for a virtual machine whose operations include construction of types, terms, and the primitive inference rules of higher order logic.
+An article is a concrete representation of a theory package, consisting of instructions for a virtual machine whose operations include construction of types and terms, and the primitive inference rules of higher order logic.
 To read an article, an interactive theorem prover performs the primitive inferences and other instructions listed in the file.
@@
 The main advantage to using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems.
-For example, the numerals 3, 5 and 8 in the example goal term can be encoded in binary using the standard constants \texttt{Number.Natural.bit0} and \texttt{Number.Natural.bit1}.
+For example, the numerals $3$, $5$ and $8$ in the example goal term can be encoded in binary using the standard constants \texttt{Number.Natural.bit0} and \texttt{Number.Natural.bit1}.
 The full names and properties of these constants are indicated in the \OpenTheory standard library, and interactive theorem provers can maintain translations to and from their local names and theorems.
@@
 
-\item\label{req:ckpt} A facility for check-pointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
+\item\label{req:ckpt} A facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
 \end{enumerate}

### 2012-02-13 paper.tex
ditch 1 reference for evidential tool bus

@@
 
-The idea of separate reasoning tools communicating to enhance a proof development environment is also being pursued in the Evidential Tool Bus~\cite{DBLP:conf/icfem/Rushby05,fk:etb} and the MathServe System~\cite{DBLP:conf/cade/ZimmerA06}.
+The idea of separate reasoning tools communicating to enhance a proof development environment is also being pursued in the Evidential Tool Bus~\cite{DBLP:conf/icfem/Rushby05} and the MathServe System~\cite{DBLP:conf/cade/ZimmerA06}.
 This idea is a natural extension of the integration of automatic tools with interactive theorem provers.

### 2012-02-13 paper.tex
write a small paragraph for the conclusion

@@
 
-\section{Conclusions}
+\section{Conclusion}
 
-We have shown how, using \OpenTheory for communication, we can make tactics for higher-order logic reasoning tasks available to multiple proof assistants and independently maintainable.
-
-<Proof assistants enable the development of formal theories both by providing an environment for managing formalisations, that is, keeping track of the details and maintaning consistency, and also by providing customised tools for specific reasoning tasks.>
-
-<Tactics tailored to a particular proof system would suffer performance hits if the factorisation through the OpenTheory kernel cannot be done efficiently.>
-<The article format might admit optimisation by restructuring low-level proofs, and such optimisers could be widely applied.>
-<Tactics where proof search dominates proof replay offer the best prospects, since the search can be done elsewhere (in parallel), the overhead is mainly in transporting and replaying proofs, and queries might be cached>
-
-<Other examples of functionality that could usefully be made prover-independent include packages for defining recursive datatypes, quotient types, nominal types, or for solving classes of arithmetic problems (decision procedures).
-- worth mentioning?
-Better to only list future work one actually intends to do.
-(E.g.
-isabelle and sledgehammer)>
-
-<extensions to other logics?>
+We have shown how, using \OpenTheory for communication, we can write tools for higher order logic reasoning tasks as standalone tactics, making them available to multiple interactive theorem provers and independently maintainable.
+Existing proof tools can be extracted from their home systems for reuse.
+There is a substantial but hopefully tolerable overhead of communicating goals and proofs over the web.
 

### 2012-02-13 paper.tex
tweaks in sec 2.2, make clearer why checkpointing req is there

@@
 
-\item\label{req:ckpt} A facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
+\item\label{req:ckpt} A facility for making a standalone executable encompassing the tactic functionality without the usual interface to the interactive theorem prover.
 \end{enumerate}
 
-Just as the requirements of an interactive theorem prover to use standalone tactics also enable it to import \OpenTheory packages in general, the requirements to create standalone tactics (specifically Requirements~\ref{req:io} and~\ref{req:log}) also enable a system to create and export \OpenTheory packages.
+Just as the requirements of an interactive theorem prover to use standalone tactics also enable it to import \OpenTheory packages in general, the first two requirements to create standalone tactics also enable a system to create and export \OpenTheory packages.
+(The last requirement enables running on a server.)
 
-Requirement~\ref{req:log} above poses the highest barrier: a standalone tactic must record proofs at a level of detail sufficient to create proofs of the same theorems using the \OpenTheory kernel.
+Requirement~\ref{req:log} poses the highest barrier: a standalone tactic must record a proof at a level of detail sufficient to prove the same theorem using the \OpenTheory kernel.
 It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules, as is the case for HOL4 and HOL Light.
@@
 For each tactic, we captured a function that reads an article representing the input term, runs the tactic (recording the proof), and writes an article containing the proof.
-The situation for HOL Light is more complicated because OCaml does not provide such an `export' facility, and HOL Light does not use stored theories.
-We used a general-purpose checkpointing facility to capture a HOL Light process at the point where it is ready to read an article and run a tactic.
+The situation for HOL Light is more complicated because OCaml does not provide such an `export' facility, and a HOL Light session typically starts by proving the standard library.
+We used a general-purpose checkpointing facility to capture a HOL Light session at the point where it is ready to read an article and run a tactic.
 

### 2012-02-13 paper.tex
tweaks to sec 3 and local timing data

@@
 \begin{center}
-\begin{tabular}{||@{~}l@{~}||@{}l@{~}r@{~}|@{~}r@{~}l@{}||@{~}r@{~}||@{}l@{~}r@{~}|@{~}r@{~}l@{}||@{~}r@{~}||}
+\begin{tabular}{||@{~}l@{~}||@{}l@{~}r@{~}|@{~}r@{~}l@{}||@{~}r@{~}||@{}l@{~}r@{~}|@{~}r@{~}l@{}||@{~}r@{~}||r@{~}||}
 \hline
-\textbf{Problem} & & \multicolumn{2}{c}{\textbf{Goal}} & & \textbf{Web} & & \multicolumn{2}{c}{\textbf{Proof}} & & \textbf{Total} \\
+\textbf{Problem} & & \multicolumn{2}{c}{\textbf{Goal}} & & \textbf{Web} & & \multicolumn{2}{c}{\textbf{Proof}} & & \textbf{Total} & \textbf{Local} \\
 \cline{2-4} \cline{7-9}
-\textbf{Name} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} \\
- & & \textbf{(b)} & \textbf{(s)} & & \textbf{(s)} & & \textbf{(b)} & \textbf{(s)} & & \textbf{(s)} \\
+\textbf{Name} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} & \textbf{Time} \\
+ & & \textbf{(b)} & \textbf{(s)} & & \textbf{(s)} & & \textbf{(b)} & \textbf{(s)} & & \textbf{(s)} & \textbf{(s)} \\
 \hline
-QBF${}_1$ & & 927 & 0.001 & & 1.064 & & 10,991 & 0.022 & & 1.088 \\
-QBF${}_2$ & & 1,474 & 0.001 & & 1.892 & & 79,944 & 0.139 & & 2.034 \\
-QBF${}_3$ & & 1,546 & 0.001 & & 1.821 & & 91,639 & 0.172 & & 1.996 \\
-SKICo${}_1$ & & 927 & 0.001 & & 1.212 & & 20,047 & 0.041 & & 1.255 \\
-SKICo${}_2$ & & 1,474 & 0.002 & & 1.557 & & 52,249 & 0.113 & & 1.673 \\
-SKICo${}_3$ & & 1,546 & 0.002 & & 1.716 & & 60,642 & 0.125 & & 1.844 \\
+QBF${}_1$ & & 927 & 0.001 & & 1.064 & & 10,991 & 0.022 & & 1.088 & 0.002\\
+QBF${}_2$ & & 1,474 & 0.001 & & 1.892 & & 79,944 & 0.139 & & 2.034 & 0.024\\
+QBF${}_3$ & & 1,546 & 0.001 & & 1.821 & & 91,639 & 0.172 & & 1.996 & 0.024\\
+SKICo${}_1$ & & 927 & 0.001 & & 1.212 & & 20,047 & 0.041 & & 1.255 & 0.000\\
+SKICo${}_2$ & & 1,474 & 0.002 & & 1.557 & & 52,249 & 0.113 & & 1.673 & 0.001\\
+SKICo${}_3$ & & 1,546 & 0.002 & & 1.716 & & 60,642 & 0.125 & & 1.844 & 0.005\\
 \hline
@@
 
-\textbf{RAMANA: perhaps add a column for the time taken to execute the tactic locally and talk about it briefly in the execution time paragraph below.}
-
-Table~\ref{tab:performance} shows preliminary performance results executing the QBF and SKICo standalone tactics on the following three test goals over boolean variables:
+We collected preliminary performance data for two tactics extracted from HOL4, made standalone, and called in HOL Light.
+They are QBF~\cite{DBLP:conf/itp/KumarW11}, which proves quantified boolean formulas, and SKICo, which rewrites terms to combinatory form, for example proving $\vdash(\forall{x}.\ x\lor\lnot x)=(\forall)\ (\mathtt{S}\ (\lor)\ (\lnot))$.
+We used the following three test goals:
 \begin{enumerate}
@@
 For each invocation of a standalone tactic on a test goal, Table~\ref{tab:performance} profiles the time and space requirements of the three phases of execution: encoding the goal as an article; communicating and executing the standalone tactic via the web; and replaying the proof article received.
+For comparison, the time to run the tactic locally within HOL4 is given in the rightmost column.
 
-Table~\ref{tab:performance} shows the size of the articles for the test goals and the resulting proofs is comparable to the typical size of web requests and the resulting pages, so we can be confident that we are within the normal operating range of the web tools we use (curl on the client and CGI scripts on the server).
+Table~\ref{tab:performance} shows the size of the articles for the test goals and the resulting proofs is comparable to the typical size of web requests and the resulting pages, so we can be confident that we are within the normal operating range of the web tools we use (\texttt{curl} on the client and CGI scripts on the server).
 For standalone tactics generating large proofs we may wish to compress them using the \texttt{gzip} tool before sending them to the client---previous experiments showed a 90\% compression ratio is typical on articles~\cite{hurd2009}.
 
-As for the execution time, Table~\ref{tab:performance} shows the bulk of the time is spent communicating and executing the standalone tactic via the web, so there is not much profit to be gained in optimizing the client.
+As for the execution time, Table~\ref{tab:performance} shows the bulk of the time is spent executing the standalone tactic on the web and communicating, so there is not much to be gained from optimizing the client.
+Web time includes reading the goal article, running the proof tool, writing the proof article, and network communication.
 Using the \texttt{traceroute} tool we see a 0.173s network delay between the test client in Portland, OR, USA and the test server in Cambridge, UK, which accounts for at least 0.346s of delay.

### 2012-02-13 paper.tex
clarify last sentence of conclusion
some space-saving tweaks
(Mike's suggestions)

@@
 \]
-decides to pass it off to a standalone tactic for linear arithmetic.\footnote{This is a representation of the problem asking for a proof that \pounds 3 and \pounds 5 coins can be used to make \pounds $n$ for any natural number $n$ that is at least 8.}
+decides to pass it off to a standalone tactic for linear arithmetic.
 
@@
 
-The main advantage to using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems.
+The main benefit of using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems.
 For example, the numerals $3$, $5$ and $8$ in the example goal term can be encoded in binary using the standard constants \texttt{Number.Natural.bit0} and \texttt{Number.Natural.bit1}.
@@
 \begin{enumerate}
-\item\label{req:io} An \OpenTheory article file reader and writer.
+\item\label{req:io} Ability to read and write \OpenTheory article files.
 
-\item\label{req:log} A facility for recording proofs and converting them to the \OpenTheory logical kernel.
+\item\label{req:log} Ability to record proofs and present them over the \OpenTheory kernel.
 
-\item\label{req:ckpt} A facility for making a standalone executable encompassing the tactic functionality without the usual interface to the interactive theorem prover.
+\item\label{req:ckpt} Ability to make a standalone executable encompassing the tactic functionality without the usual interface to the interactive theorem prover.
 \end{enumerate}
@@
 Existing proof tools can be extracted from their home systems for reuse.
-There is a substantial but hopefully tolerable overhead of communicating goals and proofs over the web.
+There is a substantial but hopefully tolerable overhead of communicating goals and proofs in article files over the web.
 

### 2012-02-13 paper.tex
tweak for space and grammar fix
various tweaks

@@
 Proof tools in interactive theorem provers are usually developed within and tied to a specific system, which leads to a duplication of effort to make the functionality available in different systems.
-Many verification projects would benefit from access to the union of all proof tools implemented across interactive theorem provers.
-Using \OpenTheory as a language for communicating between systems, we show how to turn an existing proof tool implemented for one system into a standalone tactic available to many systems via the web.
-This enables, for example, LCF-style proof reconstruction efforts for integrated external proof tools to be shared by users of different interactive theorem provers and removes the need for each user to install the external proof tool.
+Many verification projects would benefit from access to proof tools developed in other systems.
+Using \OpenTheory as a language for communicating between systems, we show how to turn a proof tool implemented for one system into a standalone tactic available to many systems via the web.
+This enables, for example, LCF-style proof reconstruction efforts to be shared by users of different interactive theorem provers and removes the need for each user to install the external tool being integrated.
 \end{abstract}
@@
 For instance, the Metis first-order prover~\cite{hurd2003d} is integrated in at least three HOL-based provers, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
-Another example is Slind's TFL package for defining recursive functions~\cite{slindphd}, originally implemented for both Isabelle/HOL and HOL4, and recently superceded in Isabelle/HOL by Krauss's function definition package~\cite{DBLP:journals/jar/Krauss10}.
+Slind's TFL package for defining recursive functions~\cite{slindphd}, originally implemented for both Isabelle/HOL and HOL4, was superseded in Isabelle/HOL by Krauss's function definition package~\cite{DBLP:journals/jar/Krauss10}.
 The improvements of Krauss's method over TFL ought to be applicable to other HOL-based provers, but a direct reimplementation would require substantial effort.
@@
 
-As an example, suppose the user of an interactive theorem prover is faced with the goal
+An example: the user of an interactive theorem prover faced with the goal
 \[
@@
 An article is a concrete representation of a theory package, consisting of instructions for a virtual machine whose operations include construction of types and terms, and the primitive inference rules of higher order logic.
-To read an article, an interactive theorem prover performs the primitive inferences and other instructions listed in the file.
+To \emph{read} an article, an interactive theorem prover performs the primitive inferences and other instructions listed in the file.
 The \OpenTheory logical kernel is based on HOL Light's logical kernel, and the instructions are chosen to make it easy to read articles into any system that can prove theorems of higher order logic.
@@
 By taking $\Delta$ to be the set of theorems proved by a proof tool and $\Gamma$ to be the set of theorems used by the proof tool, we can view the result of executing a proof tool as a logical theory.
-In our example above of using a linear arithmetic standalone tactic implementing on the given goal, this theory might be
+In our example above of using a linear arithmetic standalone tactic on the given goal, this theory might be
 \[
@@
 \end{enumerate}
-For each invocation of a standalone tactic on a test goal, Table~\ref{tab:performance} profiles the time and space requirements of the three phases of execution: encoding the goal as an article; communicating and executing the standalone tactic via the web; and replaying the proof article received.
+For each invocation of a standalone tactic on a test goal, Table~\ref{tab:performance} profiles the time and space requirements of three phases of execution: encoding the goal as an article; communicating and executing the standalone tactic via the web; and replaying the proof article received.
 For comparison, the time to run the tactic locally within HOL4 is given in the rightmost column.
 
-Table~\ref{tab:performance} shows the size of the articles for the test goals and the resulting proofs is comparable to the typical size of web requests and the resulting pages, so we can be confident that we are within the normal operating range of the web tools we use (\texttt{curl} on the client and CGI scripts on the server).
-For standalone tactics generating large proofs we may wish to compress them using the \texttt{gzip} tool before sending them to the client---previous experiments showed a 90\% compression ratio is typical on articles~\cite{hurd2009}.
+The sizes of the articles for the test goals and the resulting proofs are comparable to the typical size of web requests and the resulting pages, so we can be confident that we are within the normal operating range of the web tools we use (\texttt{curl} on the client and CGI scripts on the server).
+For standalone tactics generating large proofs we may wish to compress them using \texttt{gzip} before sending them to the client---previous experiments showed a 90\% compression ratio is typical on articles~\cite{hurd2009}.
 
-As for the execution time, Table~\ref{tab:performance} shows the bulk of the time is spent executing the standalone tactic on the web and communicating, so there is not much to be gained from optimizing the client.
-Web time includes reading the goal article, running the proof tool, writing the proof article, and network communication.
-Using the \texttt{traceroute} tool we see a 0.173s network delay between the test client in Portland, OR, USA and the test server in Cambridge, UK, which accounts for at least 0.346s of delay.
+As for the execution time, the bulk of the time is spent executing the standalone tactic on the web and communicating, so there is not much to be gained from optimizing the client.
+``Web Time'' includes reading the goal article, running the proof tool, writing the proof article, and network communication.
+Using \texttt{traceroute} we see a 0.173s network delay between the test client in Portland, OR, USA and the test server in Cambridge, UK, which accounts for at least 0.346s of delay.
 The overall time is in the 1--2s range, which is very slow for workaday tactics but may well be tolerated by the user to gain access to the functionality of a proof tool on another system.
@@
 
-An impressive example of providing reasoning infrastructure over the web is the System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}.
+An impressive example of providing reasoning infrastructure over the web is System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}.
 Here, multiple automatic theorem provers can be called on a problem (\eg satisfiability of a formula, or whether a goal follows from assumptions) expressed in the standard TPTP format.

### 2012-02-13 paper.tex
more tweaks, now under 6 pages!

@@
 Examples of duplicated effort on tactics include the integration of external tools into HOL-based provers.
-For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into a higher-order logic based system (HOL4 in one case and HOL Light in the other).
-Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems (HOL4, HOL Light, and Isabelle/HOL); each integration requires a separate implementation.
+For instance, Kumar and Weber~\cite{DBLP:conf/itp/KumarW11} and Kun\v{c}ar~\cite{DBLP:conf/itp/Kuncar11} give independent integrations of a quantified boolean formula solver into HOL4 and HOL Light.
+Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems; each integration requires a separate implementation.
 Sledgehammer~\cite{DBLP:conf/frocos/BlanchetteBN11} is only available for Isabelle/HOL, but would be useful if available for other provers.
@@
 The input for the standalone tactic is the goal, and the output is a proof of the theorem.
-Standalone tactics use the \OpenTheory article format for communicating both goal terms and proofs of theorems.
+Standalone tactics use the \OpenTheory article format for communicating terms and proofs.
 The interactive theorem prover serializes the goal term from its local internal format to an article file, and sends the article over the web to the standalone tactic.

### 2012-02-13 paper.tex
some more minor tweaks

@@
 The \OpenTheory standard library~\cite{hurd2011} formalises this shared base as a collection of theory packages containing proofs written in the simple \emph{article} format designed for storing and sharing higher order logic theories~\cite{hurd2009}.
-We use \OpenTheory as a language for users of interactive theorem provers to communicate with proof tools on a remote server, and thereby obtain the following two benefits:
+We use \OpenTheory as a language for interactive theorem provers to communicate with proof tools on a remote server, and thereby obtain the following two benefits:
 \begin{enumerate}
@@
 
-Requirement~\ref{req:log} poses the highest barrier: a standalone tactic must record a proof at a level of detail sufficient to prove the same theorem using the \OpenTheory kernel.
+Requirement~\ref{req:log} poses the highest barrier: a standalone tactic must record each proof at a level of detail sufficient to prove the same theorem using the \OpenTheory kernel.
 It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules, as is the case for HOL4 and HOL Light.
@@
 The sizes of the articles for the test goals and the resulting proofs are comparable to the typical size of web requests and the resulting pages, so we can be confident that we are within the normal operating range of the web tools we use (\texttt{curl} on the client and CGI scripts on the server).
-For standalone tactics generating large proofs we may wish to compress them using \texttt{gzip} before sending them to the client---previous experiments showed a 90\% compression ratio is typical on articles~\cite{hurd2009}.
+For problems involving larger articles (long proofs or big goals) we may wish to compress them using \texttt{gzip} before sending them to the client---previous experiments showed a 90\% compression ratio is typical on articles~\cite{hurd2009}.
 
-As for the execution time, the bulk of the time is spent executing the standalone tactic on the web and communicating, so there is not much to be gained from optimizing the client.
+Turning now to execution time, the bulk of the time is spent executing the standalone tactic on the web and communicating, so there is not much to be gained from optimizing the client.
 ``Web Time'' includes reading the goal article, running the proof tool, writing the proof article, and network communication.

### 2012-02-14 paper.tex
Joe's pass

@@
 Weber and Amjad~\cite{weber09efficiently} give high-performance integrations of SAT solvers into three HOL-based systems; each integration requires a separate implementation.
-Sledgehammer~\cite{DBLP:conf/frocos/BlanchetteBN11} is only available for Isabelle/HOL, but would be useful if available for other provers.
+Sledgehammer~\cite{DBLP:conf/frocos/BlanchetteBN11} is only available for Isabelle/HOL, but its functionality would also be useful in other systems.
 
-Besides duplicate initial effort, costs of maintenance can be multiplied over different provers, and improvements in functionality can become restricted to a single prover unnecessarily.
-For instance, the Metis first-order prover~\cite{hurd2003d} is integrated in at least three HOL-based provers, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
+In addition to the development effort, the costs of maintenance can also be multiplied over different systems, and improvements in functionality can become restricted to a single system unnecessarily.
+For instance, the Metis first-order prover~\cite{hurd2003d} is integrated in multiple systems in the HOL family, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
 Slind's TFL package for defining recursive functions~\cite{slindphd}, originally implemented for both Isabelle/HOL and HOL4, was superseded in Isabelle/HOL by Krauss's function definition package~\cite{DBLP:journals/jar/Krauss10}.
@@
 \begin{enumerate}
-\item\textbf{Proof Articles:} A standard format to encode the goals that will be sent to the proof tools and the proofs that will be received.
+\item\textbf{Proof Articles:} A standard format to encode the goals that will be sent to the remote proof tools and the proofs that will be received in response.
 
@@
 
-The input for the standalone tactic is the goal, and the output is a proof of the theorem.
+The input for the standalone tactic is the goal term, and the output is a proof of the theorem.
 Standalone tactics use the \OpenTheory article format for communicating terms and proofs.
@@
 \vdash\forall n.\ n + 0 = n \\
-\vdash\forall m,n.\ m + n = n + m \\
+\vdash\forall m,n.\ mn = nm \\
 \cdots
 \end{array}\right\}}
-{\bigl\{\vdash\forall n.\ 8\le n\Rightarrow\exists s,t.\ n = 3s + 5t\bigr\}}
+{\bigl\{\,\vdash\forall n.\ 8\le n\Rightarrow\exists s,t.\ n = 3s + 5t\,\bigr\}}
 \]
@@
 
-The main benefit of using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems.
+The main benefit of using \OpenTheory for communication is that it provides a standard ontology for fixing the meanings of constants and type operators between different systems.
 For example, the numerals $3$, $5$ and $8$ in the example goal term can be encoded in binary using the standard constants \texttt{Number.Natural.bit0} and \texttt{Number.Natural.bit1}.
@@
 
-Implementing an \OpenTheory interface to satisfy Requirements~\ref{req:artr} and~\ref{req:artw} above carries the additional benefit of giving the interactive theorem prover access to all theories stored as \OpenTheory packages, not just those that are the output of standalone tactics.
+Implementing an \OpenTheory interface to satisfy Requirements~\ref{req:artr} and~\ref{req:artw} above carries the additional benefit of giving the interactive theorem prover access to all logical theories stored as \OpenTheory packages, not just those that are the output of standalone tactics.
 
@@
 
-\item\label{req:log} Ability to record proofs and present them over the \OpenTheory kernel.
+\item\label{req:log} Ability to record proofs and reduce them to the \OpenTheory kernel.
 
-\item\label{req:ckpt} Ability to make a standalone executable encompassing the tactic functionality without the usual interface to the interactive theorem prover.
+\item\label{req:ckpt} Ability to make a standalone executable encompassing the tactic functionality separated from the usual interface to the interactive theorem prover.
 \end{enumerate}
@@
 Requirement~\ref{req:log} poses the highest barrier: a standalone tactic must record each proof at a level of detail sufficient to prove the same theorem using the \OpenTheory kernel.
-It helps if the system containing the tactic uses the LCF architecture, that is, theorems can only be created by a small number of primitive inference rules, as is the case for HOL4 and HOL Light.
-
-We met Requirement~\ref{req:log} by augmenting the internal theorem type with a type of proofs, to track primitive inferences used.
-Then we expressed each primitive inference as a derived rule above the \OpenTheory kernel.
-In many cases, there is a direct counterpart (\eg reflexivity of equality is primitive in all three kernels).
-But, for example, HOL4's primitive rule for definition by specification must be emulated in \OpenTheory, for example by using Hilbert choice.
+The following method can be used if the system implementing the tactic has an LCF-style design, that is, theorems can only be created by a small number of primitive inference rules: (i) augment the internal theorem type with a type of proofs to track the system primitive inferences used; and (ii) express each system primitive inference as a derived rule of the \OpenTheory kernel.
+We applied this method to meet Requirement~\ref{req:log} for both HOL4 and HOL Light.
+For some primitive inferences (such as reflexivity of equality) there is a direct translation to the \OpenTheory logical kernel.
+But, for example, HOL4's primitive rule for definition by specification must be emulated in the \OpenTheory kernel, for example by using Hilbert choice.
 Although Isabelle/HOL uses the LCF architecture, its logical kernel is quite different from \OpenTheory; we therefore expect translating Isabelle/HOL proofs to be more difficult than HOL4 and HOL Light proofs.
 
-To satisfy Requirement~\ref{req:ckpt} in HOL4, we used the `export' facility of the PolyML compiler, which creates a standalone executable that runs an ML function.
+To satisfy Requirement~\ref{req:ckpt} in HOL4, we used the `export' facility of the Poly/ML compiler, which creates a standalone executable that runs an ML function.
 For each tactic, we captured a function that reads an article representing the input term, runs the tactic (recording the proof), and writes an article containing the proof.
@@
 
-We collected preliminary performance data for two tactics extracted from HOL4, made standalone, and called in HOL Light.
-They are QBF~\cite{DBLP:conf/itp/KumarW11}, which proves quantified boolean formulas, and SKICo, which rewrites terms to combinatory form, for example proving $\vdash(\forall{x}.\ x\lor\lnot x)=(\forall)\ (\mathtt{S}\ (\lor)\ (\lnot))$.
+We collected preliminary performance data for two test standalone tactics extracted from HOL4, and called from HOL Light.
+They are QBF~\cite{DBLP:conf/itp/KumarW11}, which proves quantified boolean formulas, and SKICo, which rewrites terms to combinatory form, for example proving $\vdash(\forall{x}.\ x\,\lor\,\lnot x)=(\forall)\ (\mathtt{S}\ (\lor)\ (\lnot))$.
 We used the following three test goals:
@@
 \end{enumerate}
-For each invocation of a standalone tactic on a test goal, Table~\ref{tab:performance} profiles the time and space requirements of three phases of execution: encoding the goal as an article; communicating and executing the standalone tactic via the web; and replaying the proof article received.
+For each invocation of a standalone tactic on a test goal, Table~\ref{tab:performance} profiles the time and space requirements of the three phases of execution: encoding the goal as an article; communicating with and executing the standalone tactic via the web; and replaying the proof article received.
 For comparison, the time to run the tactic locally within HOL4 is given in the rightmost column.
@@
 The sizes of the articles for the test goals and the resulting proofs are comparable to the typical size of web requests and the resulting pages, so we can be confident that we are within the normal operating range of the web tools we use (\texttt{curl} on the client and CGI scripts on the server).
-For problems involving larger articles (long proofs or big goals) we may wish to compress them using \texttt{gzip} before sending them to the client---previous experiments showed a 90\% compression ratio is typical on articles~\cite{hurd2009}.
+For problems involving larger articles (bigger goals or longer proofs) we may wish to compress them using \texttt{gzip} before sending them over the network---previous experiments showed a compression ratio of 90\% is typical for article files~\cite{hurd2009}.
 
-Turning now to execution time, the bulk of the time is spent executing the standalone tactic on the web and communicating, so there is not much to be gained from optimizing the client.
-``Web Time'' includes reading the goal article, running the proof tool, writing the proof article, and network communication.
+Turning now to execution time, we can see that it is significantly more expensive to call a standalone tactic over the web compared to executing it locally.
+However, most of the time is spent on `Web Time', which includes communicating with the remote standalone tactic and waiting for it to read the goal article, run the proof tool, and write the resulting proof article.
 Using \texttt{traceroute} we see a 0.173s network delay between the test client in Portland, OR, USA and the test server in Cambridge, UK, which accounts for at least 0.346s of delay.
@@
 
-An impressive example of providing reasoning infrastructure over the web is System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}.
-Here, multiple automatic theorem provers can be called on a problem (\eg satisfiability of a formula, or whether a goal follows from assumptions) expressed in the standard TPTP format.
-The usual scenario is a decision problem in first-order logic, but TPTP includes support for higher-order terms (THF format) and for expressing proofs in a standard language (TSTP).
+An impressive example of providing reasoning infrastructure over the web is System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}, which enables a user to remotely execute a collection of automatic theorem provers on a problem expressed in a standard format.
+The usual scenario is to decide the validity of first-order logic formula (TPTP format), but there is also support for higher-order terms (THF format), and for returning proofs expressed in a standard language (TSTP format).
 

### 2012-02-14 paper.tex
Tweaks

@@
 
-There are many LCF-style systems for interactively developing machine-checked formal theories, including HOL4~\cite{slind08brief}, HOL Light~\cite{DBLP:conf/tphol/Harrison09a}, ProofPower~\cite{ProofPower}, and Isabelle/HOL~\cite{wenzel08isabelle}.
+There are many LCF-style systems for interactively developing machine-checked formal theories, including HOL4~\cite{slind08brief}, HOL Light~\cite{DBLP:conf/tphol/Harrison09a}, ProofPower~\cite{ProofPower} and Isabelle/HOL~\cite{wenzel08isabelle}.
 The logic implemented by these systems is essentially the same, but the collections of theory libraries and proof tools built on top of the logical kernels differ.
@@
 In addition to the development effort, the costs of maintenance can also be multiplied over different systems, and improvements in functionality can become restricted to a single system unnecessarily.
-For instance, the Metis first-order prover~\cite{hurd2003d} is integrated in multiple systems in the HOL family, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
+For instance, the Metis first order logic prover~\cite{hurd2003d} is integrated in multiple systems in the HOL family, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
 Slind's TFL package for defining recursive functions~\cite{slindphd}, originally implemented for both Isabelle/HOL and HOL4, was superseded in Isabelle/HOL by Krauss's function definition package~\cite{DBLP:journals/jar/Krauss10}.
@@
 
-% Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem.\footnote{For example, we could represent $t$ by the theory $\emptyset\rhd\{\vdash{t=t}\}$.
+% Furthermore, we can represent a higher order logic term in a theory by including the term in the statement of a theorem.\footnote{For example, we could represent $t$ by the theory $\emptyset\rhd\{\vdash{t=t}\}$.
 % (The article format allows for multiple occurrences of $t$ to be shared.)
@@
 An impressive example of providing reasoning infrastructure over the web is System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}, which enables a user to remotely execute a collection of automatic theorem provers on a problem expressed in a standard format.
-The usual scenario is to decide the validity of first-order logic formula (TPTP format), but there is also support for higher-order terms (THF format), and for returning proofs expressed in a standard language (TSTP format).
+The usual scenario is to decide the validity of first order logic formulas (TPTP format), but there is also support for higher order terms (THF format), and for returning proofs expressed in a standard language (TSTP format).
 

### 2012-02-14 paper.tex
Clarification

@@
 \hline
-\textbf{Problem} & & \multicolumn{2}{c}{\textbf{Goal}} & & \textbf{Web} & & \multicolumn{2}{c}{\textbf{Proof}} & & \textbf{Total} & \textbf{Local} \\
+\textbf{Tactic-} & & \multicolumn{2}{c}{\textbf{Goal}} & & \textbf{Web} & & \multicolumn{2}{c}{\textbf{Proof}} & & \textbf{Total} & \textbf{Local} \\
 \cline{2-4} \cline{7-9}
-\textbf{Name} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} & \textbf{Time} \\
+\textbf{Problem} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} & & \textbf{Size} & \textbf{Time} & & \textbf{Time} & \textbf{Time} \\
  & & \textbf{(b)} & \textbf{(s)} & & \textbf{(s)} & & \textbf{(b)} & \textbf{(s)} & & \textbf{(s)} & \textbf{(s)} \\
 \hline
-QBF${}_1$ & & 927 & 0.001 & & 1.064 & & 10,991 & 0.022 & & 1.088 & 0.002\\
-QBF${}_2$ & & 1,474 & 0.001 & & 1.892 & & 79,944 & 0.139 & & 2.034 & 0.024\\
-QBF${}_3$ & & 1,546 & 0.001 & & 1.821 & & 91,639 & 0.172 & & 1.996 & 0.024\\
-SKICo${}_1$ & & 927 & 0.001 & & 1.212 & & 20,047 & 0.041 & & 1.255 & 0.000\\
-SKICo${}_2$ & & 1,474 & 0.002 & & 1.557 & & 52,249 & 0.113 & & 1.673 & 0.001\\
-SKICo${}_3$ & & 1,546 & 0.002 & & 1.716 & & 60,642 & 0.125 & & 1.844 & 0.005\\
+QBF-1 & & 927 & 0.001 & & 1.064 & & 10,991 & 0.022 & & 1.088 & 0.002\\
+QBF-2 & & 1,474 & 0.001 & & 1.892 & & 79,944 & 0.139 & & 2.034 & 0.024\\
+QBF-3 & & 1,546 & 0.001 & & 1.821 & & 91,639 & 0.172 & & 1.996 & 0.024\\
+SKICo-1 & & 927 & 0.001 & & 1.212 & & 20,047 & 0.041 & & 1.255 & 0.000\\
+SKICo-2 & & 1,474 & 0.002 & & 1.557 & & 52,249 & 0.113 & & 1.673 & 0.001\\
+SKICo-3 & & 1,546 & 0.002 & & 1.716 & & 60,642 & 0.125 & & 1.844 & 0.005\\
 \hline

### 2012-02-14 paper.tex
Final tweaks

@@
 We collected preliminary performance data for two test standalone tactics extracted from HOL4, and called from HOL Light.
-They are QBF~\cite{DBLP:conf/itp/KumarW11}, which proves quantified boolean formulas, and SKICo, which rewrites terms to combinatory form, for example proving $\vdash(\forall{x}.\ x\,\lor\,\lnot x)=(\forall)\ (\mathtt{S}\ (\lor)\ (\lnot))$.
+They are QBF~\cite{DBLP:conf/itp/KumarW11}, which proves quantified boolean formulas, and SKICo, which rewrites terms to combinatory form like so: $\vdash(\forall{x}.\ x\lor\lnot x)=(\forall)\ (\mathtt{S}\ (\lor)\ (\lnot))$.
 We used the following three test goals:

### 2012-05-12 paper.tex
Added a reference to PROSPER
add newlines between institutions and emails
in accordance with Springer guidelines
Minor changes and a paragraph on the PROSPER related work

@@
 
-\institute{University of Cambridge, \email{Ramana.Kumar@cl.cam.ac.uk}\and Galois, Inc., \email{joe@gilith.com}}
+\institute{ University of Cambridge\\\email{Ramana.Kumar@cl.cam.ac.uk}\\\and Galois, Inc.\\\email{joe@gilith.com}}
 
@@
 
-In addition to the development effort, the costs of maintenance can also be multiplied over different systems, and improvements in functionality can become restricted to a single system unnecessarily.
+In addition to the development effort, the cost of maintenance can also be multiplied over different systems, and improvements in functionality can become restricted to a single system unnecessarily.
 For instance, the Metis first order logic prover~\cite{hurd2003d} is integrated in multiple systems in the HOL family, but the HOL4 version is very old compared to the latest version in Isabelle/HOL.
@@
 The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.
-For tactic extraction to succeed, the key requirements of the interactive theorem prover are:
+For tactic extraction to succeed, the key requirements for the interactive theorem prover are:
 \begin{enumerate}
@@
 
+The PROSPER project~\cite{dennis2000} pioneered the technique of sending goals over the internet to a remote solver, and packaging such procedures as tactics in the HOL4 theorem prover.
+The standalone tactics described in this paper further systematize this by using \OpenTheory as a standard language and ontology to make it easier for interactive theorem provers and remote solvers to communicate.
+
 An impressive example of providing reasoning infrastructure over the web is System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}, which enables a user to remotely execute a collection of automatic theorem provers on a problem expressed in a standard format.

### 2012-05-12 paper.tex
change references to 'web' to 'internet'

@@
 Many verification projects would benefit from access to proof tools developed in other systems.
-Using \OpenTheory as a language for communicating between systems, we show how to turn a proof tool implemented for one system into a standalone tactic available to many systems via the web.
+Using \OpenTheory as a language for communicating between systems, we show how to turn a proof tool implemented for one system into a standalone tactic available to many systems via the internet.
 This enables, for example, LCF-style proof reconstruction efforts to be shared by users of different interactive theorem provers and removes the need for each user to install the external tool being integrated.
@@
 Standalone tactics use the \OpenTheory article format for communicating terms and proofs.
-The interactive theorem prover serializes the goal term from its local internal format to an article file, and sends the article over the web to the standalone tactic.
+The interactive theorem prover serializes the goal term from its local internal format to an article file, and sends the article over the internet to the standalone tactic.
 If successful, the standalone tactic sends back another article file encoding a proof of the goal, which the interactive theorem prover replays through its logical kernel to create the desired theorem.
@@
 
-\item\label{req:comm} Ability to communicate over the web with external programs.
+\item\label{req:comm} Ability to communicate with external programs.
 \end{enumerate}
@@
 \hline
-\textbf{Tactic-} & & \multicolumn{2}{c}{\textbf{Goal}} & & \textbf{Web} & & \multicolumn{2}{c}{\textbf{Proof}} & & \textbf{Total} & \textbf{Local} \\
+\textbf{Tactic-} & & \multicolumn{2}{c}{\textbf{Goal}} & & \textbf{Remote} & & \multicolumn{2}{c}{\textbf{Proof}} & & \textbf{Total} & \textbf{Local} \\
 \cline{2-4} \cline{7-9}
@@
 \end{enumerate}
-For each invocation of a standalone tactic on a test goal, Table~\ref{tab:performance} profiles the time and space requirements of the three phases of execution: encoding the goal as an article; communicating with and executing the standalone tactic via the web; and replaying the proof article received.
+For each invocation of a standalone tactic on a test goal, Table~\ref{tab:performance} profiles the time and space requirements of the three phases of execution: encoding the goal as an article; communicating with and executing the standalone tactic remotely; and replaying the proof article received.
 For comparison, the time to run the tactic locally within HOL4 is given in the rightmost column.
@@
 
-Turning now to execution time, we can see that it is significantly more expensive to call a standalone tactic over the web compared to executing it locally.
-However, most of the time is spent on `Web Time', which includes communicating with the remote standalone tactic and waiting for it to read the goal article, run the proof tool, and write the resulting proof article.
+Turning now to execution time, we can see that it is significantly more expensive to call a standalone tactic over the internet compared to executing it locally.
+However, most of the time is spent on `Remote Time', which includes communicating with the remote standalone tactic and waiting for it to read the goal article, run the proof tool, and write the resulting proof article.
 Using \texttt{traceroute} we see a 0.173s network delay between the test client in Portland, OR, USA and the test server in Cambridge, UK, which accounts for at least 0.346s of delay.
@@
 
-An impressive example of providing reasoning infrastructure over the web is System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}, which enables a user to remotely execute a collection of automatic theorem provers on a problem expressed in a standard format.
+An impressive example of providing reasoning infrastructure over the internet is System on TPTP~\cite{DBLP:conf/lpar/Sutcliffe10}, which enables a user to remotely execute a collection of automatic theorem provers on a problem expressed in a standard format.
 The usual scenario is to decide the validity of first order logic formulas (TPTP format), but there is also support for higher order terms (THF format), and for returning proofs expressed in a standard language (TSTP format).
@@
 Existing proof tools can be extracted from their home systems for reuse.
-There is a substantial but hopefully tolerable overhead of communicating goals and proofs in article files over the web.
+There is a substantial but hopefully tolerable overhead of communicating goals and proofs in article files over the internet.
 

### 2012-05-12 paper.tex
tweak some "requirements of/for/on"

@@
 The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.
-For tactic extraction to succeed, the key requirements for the interactive theorem prover are:
+For tactic extraction to succeed, the key requirements on the interactive theorem prover are:
 \begin{enumerate}
@@
 
-Just as the requirements of an interactive theorem prover to use standalone tactics also enable it to import \OpenTheory packages in general, the first two requirements to create standalone tactics also enable a system to create and export \OpenTheory packages.
+Just as the requirements for an interactive theorem prover to use standalone tactics also enable it to import \OpenTheory packages in general, the first two requirements to create standalone tactics also enable a system to create and export \OpenTheory packages.
 (The last requirement enables running on a server.)
